---
title: "Feasts & fables: time series analysis using R"
author: "Rob J Hyndman"
date: "FRB 28 Sep 2021"
toc: true
classoption: aspectratio=169
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 8
    fig_height: 3.3
    keep_tex: no
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = TRUE, warning = FALSE, cache = TRUE,
  dev.args = list(pointsize = 9)
)
```

```{r packages, echo=FALSE, include=FALSE}
# Load packages
library(fpp3)
library(patchwork)
library(glue)
library(stringr)

# Colors to be viridis for continuous scales and Okabe for discrete scales
options(
  digits = 3,
  width = 85,
  ggplot2.continuous.color = "viridis",
  ggplot2.continuous.fill = "viridis",
  ggplot2.discrete.color = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442"),
  ggplot2.discrete.fill = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442")
)

# Neater data sets
global_economy <- global_economy %>%
  select(Year, Country, GDP, Imports, Exports, Population)
tourism <- tsibble::tourism %>%
  mutate(
    State = recode(State,
      "Australian Capital Territory" = "ACT",
      "New South Wales" = "NSW",
      "Northern Territory" = "NT",
      "Queensland" = "QLD",
      "South Australia" = "SA",
      "Tasmania" = "TAS",
      "Victoria" = "VIC",
      "Western Australia" = "WA"
    ),
    Region = stringr::str_remove(Region, "Australia's ")
  )
```

# tsibble: Time series data

## Tidyverts packages

\begin{textblock}{3.8}(11,0)\begin{alertblock}{}\Large\textbf{tidyverts.org}\end{alertblock}\end{textblock}

\placefig{3}{1.4}{width=2.5cm}{tsibble.png}
\placefig{5.5}{1.4}{width=2.5cm}{tsibbledata.png}
\placefig{8}{1.4}{width=2.5cm}{feasts.png}
\placefig{10.5}{1.4}{width=2.5cm}{fable.png}

\pause\vspace*{2.9cm}\fontsize{7}{8}\sf

```r
library(fpp3)
```

\vspace*{-0.4cm}
\begin{alltt}
\input{startup.txt}
\end{alltt}


## `tsibble` objects

\fontsize{10}{11.2}\sf

```{r, echo = TRUE}
global_economy
```

\only<2->{\begin{textblock}{.75}(2.15,3.77)
\begin{alertblock}{}\fontsize{9}{9}\sf Index\phantom{dg}\end{alertblock}
\end{textblock}}
\only<3->{\begin{textblock}{1.6}(3.28,3.77)
\begin{alertblock}{}\fontsize{9}{9}\sf Key\phantom{dg}\end{alertblock}
\end{textblock}}
\only<4>{\begin{textblock}{6.7}(5.5,3.77)
\begin{alertblock}{}\fontsize{9}{9}\sf Measured variables\phantom{dg}\end{alertblock}
\end{textblock}}

## Australian tourism regions

```{r ausmap, echo=FALSE, message=FALSE, width=20, height=18, out.height="90%"}
library(sf)
# Use Okabe-Ito color-blind friendly color palette
state_colors <- c(
  `New South Wales` = "#56b4e9",
  `Victoria` = "#0072b2",
  `Queensland` = "#009e73",
  `South Australia` = "#f0e442",
  `Northern Territory` = "#d55e00",
  `Western Australia` = "#e69f00",
  `Tasmania` = "#cc79a7",
  `Australian Capital Territory` = "#cccccc"
)
read_sf("tourism/Tourism_Regions_2020.shp") %>%
  rename(State = "STE_NAME16") %>%
  ggplot() +
  geom_sf(aes(fill = State), alpha = 0.8) +
  theme_void() +
  scale_fill_manual(values = state_colors)
```

\only<2>{\begin{textblock}{6.5}(9.1,1.4)
\begin{block}{}\fontsize{12}{12}\sf
  \begin{itemize}\itemsep=0cm\parskip=0cm
    \item Quarterly data on visitor nights: 1998 -- 2017
    \item From \textit{National Visitor Survey}, interviews of 120,000 Australians aged 15+.
    \item Geographical hierarchy split by
    \begin{itemize}
    \item 8 states and territories
    \item 76 regions
    \end{itemize}
    \item Purpose:
    \begin{itemize}
    \item Holidays
    \item Business
    \item Visiting friends \& relatives
    \item Other
    \end{itemize}
  \end{itemize}
\end{block}
\end{textblock}}

## `tsibble` objects

\fontsize{10}{11.3}\sf

```{r, echo = TRUE}
tourism
```

\only<2->{\begin{textblock}{1.1}(2.1,3.7)
\begin{alertblock}{}\fontsize{10}{10}\sf Index\phantom{dg}\end{alertblock}
\end{textblock}}
\only<3->{\begin{textblock}{3.9}(3.65,3.7)
\begin{alertblock}{}\fontsize{10}{10}\sf Keys\phantom{dg}\end{alertblock}
\end{textblock}}
\only<4>{\begin{textblock}{1.5}(7.95,3.7)
\begin{alertblock}{}\fontsize{10}{10}\sf Measure\phantom{dg}\end{alertblock}
\end{textblock}}

\begin{textblock}{3}(9,5)
\begin{block}{}\fontsize{10}{10}\sf Domestic visitor nights in thousands by state/region and purpose.\phantom{dg}\end{block}
\end{textblock}

## `tsibble` objects

* A `tsibble` allows storage and manipulation of multiple time series in R.

* It contains:

  + An index: time information about the observation
  + Measured variable(s): numbers of interest
  + Key variable(s): optional unique identifiers for each series

* It works with tidyverse functions.

## Australian prison population

\centerline{\includegraphics[height=8cm]{Beechworth_prison}}

## Read a csv file and convert to a tsibble {-}
\fontsize{10}{11}\sf

```{r prison, message=FALSE}
prison <- readr::read_csv("data/prison_population.csv")
```
```{r prison2a, dependson="prison", echo=FALSE, message=FALSE}
prison
```
## Read a csv file and convert to a tsibble {-}
\fontsize{10}{11}\sf

```{r prison3, message=FALSE}
prison <- readr::read_csv("data/prison_population.csv") %>%
  mutate(Quarter = yearquarter(date))
```

```{r prison3a, dependson="prison3", echo=FALSE, message=FALSE}
prison
```

## Read a csv file and convert to a tsibble {-}
\fontsize{10}{11}\sf

```{r prison4, message=FALSE}
prison <- readr::read_csv("data/prison_population.csv") %>%
  mutate(Quarter = yearquarter(date)) %>%
  select(-date)
```

```{r prison4a, dependson="prison4", echo=FALSE, message=FALSE}
prison
```

## Read a csv file and convert to a tsibble {-}
\fontsize{10}{11}\sf

```{r prison5, message=FALSE}
prison <- readr::read_csv("data/prison_population.csv") %>%
  mutate(Quarter = yearquarter(date)) %>%
  select(-date) %>%
  as_tsibble(
    index = Quarter,
    key = c(state, gender, legal, indigenous)
  )
```

```{r prison5a, dependson="prison5", echo=FALSE, message=FALSE}
prison
```

## The `tsibble` index

Common time index variables can be created with these functions:

###
```{r tstable2, echo=FALSE}
tribble(
  ~`Frequency`, ~Function,
  "Annual", "`start:end`",
  "Quarterly", "`yearquarter()`",
  "Monthly", "`yearmonth()`",
  "Weekly", "`yearweek()`",
  "Daily", "`as_date()`, `ymd()`",
  "Sub-daily", "`as_datetime()`"
) %>%
  knitr::kable(booktabs = TRUE)
```


# feasts: Data visualization

## Australian holidays
\fontsize{9}{10}\sf

```{r tourism}
tourism
```

## Australian holidays
\fontsize{9}{10}\sf

```{r holidays}
holidays <- tourism %>%
  filter(Purpose == "Holiday") %>%
  group_by(State) %>%
  summarise(Trips = sum(Trips))
```

```{r, echo=FALSE}
holidays
```

## Australian holidays
\fontsize{9}{10}\sf

```{r holidays-plot, echo=TRUE, dependson="holidays"}
holidays %>% autoplot(Trips) +
  ylab("thousands of trips") +
  ggtitle("Australian domestic holiday nights")
```

## Seasonal plots
\fontsize{9}{10}\sf

```{r graphics1, fig.width=6, fig.height=6, out.width="40%"}
holidays %>% gg_season(Trips) +
  ylab("thousands of trips") +
  ggtitle("Australian domestic holiday nights")
```

## Seasonal subseries plots
\fontsize{9}{10}\sf

```{r graphics2, fig.width=9, fig.height=5, out.height="70%"}
holidays %>% gg_subseries(Trips) +
  ylab("thousands of trips") +
  ggtitle("Australian domestic holiday nights")
```

## Victorian electricity demand
\fontsize{10}{11}\sf

```{r}
vic_elec
```

## Multiple seasonal periods
\fontsize{12}{13}\sf

```{r, dev = "png", dpi = 180}
vic_elec %>% gg_season(Demand)
```

## Multiple seasonal periods
\fontsize{12}{13}\sf

```{r, dev = "png", dpi = 180}
vic_elec %>% gg_season(Demand, period = "week")
```

## Multiple seasonal periods
\fontsize{12}{13}\sf

```{r, dev = "png", dpi = 180}
vic_elec %>% gg_season(Demand, period = "day")
```

## Time plots

 * Plotted against time: `autoplot()` (each series overplotted)
 * Plotted against season: `gg_season()` (facet by key)
 * Plotted against time with seasonal facets: `gg_subseries()` (facet by key)

## Autocorrelations
\fontsize{9}{10}\sf

```{r tourismacf, warning=TRUE}
holidays %>% ACF(Trips)
```

## ACF plots
\fontsize{9}{10}\sf

```{r tourismacf2, fig.height=4, fig.width=4, out.height="70%"}
holidays %>%
  ACF(Trips) %>%
  autoplot()
```

## Google stock price
\fontsize{12}{14}\sf

```{r gafa}
gafa_stock
```

## Google stock price
\fontsize{12}{14}\sf

```{r google-2015}
google_2015 <- gafa_stock %>%
  filter(Symbol == "GOOG", year(Date) == 2015) %>%
  select(Date, Close)
google_2015
```

## Google stock price

```{r}
google_2015 %>% autoplot(Close)
```

## Google stock price
\fontsize{9}{9}\sf

```{r, message=TRUE, warning=TRUE}
google_2015 %>%
  ACF(Close, lag_max = 100)
```

## Google stock price
\fontsize{9}{9}\sf

```{r}
google_2015 %>%
  ACF(Close, lag_max = 100) %>%
  autoplot()
```

# feasts: Time series features

## Strength of seasonality and trend

\begin{alertblock}{STL decomposition}
\centerline{$y_t = T_t+S_t+R_t$}
\end{alertblock}

### Seasonal strength
$$\max\left(0, 1-\frac{\text{Var}(R_t)}{\text{Var}(S_t+R_t)}\right)$$

### Trend strength
$$\max\left(0, 1-\frac{\text{Var}(R_t)}{\text{Var}(T_t+R_t)}\right)$$

## Feature extraction and statistics

\fontsize{9}{10}\sf

```{r features}
tourism %>% features(Trips, feat_stl)
```

## Feature extraction and statistics
\fontsize{8}{9}\sf

```{r features-plot, fig.height=4, fig.width=6, out.height="65%"}
tourism %>%
  features(Trips, feat_stl) %>%
  ggplot(aes(x = trend_strength, y = seasonal_strength_year, col = Purpose)) +
  geom_point() +
  facet_wrap(vars(State))
```

\only<2->{\begin{textblock}{5.7}(8.8,6.6)
\begin{alertblock}{}\fontsize{12}{12}\sf
\begin{itemize}\tightlist
\item Holidays more seasonal than other travel.
\item WA has strongest trends.
\end{itemize}
\end{alertblock}\end{textblock}}

## Feature extraction and statistics
\fontsize{9}{9}\sf

Find the most seasonal time series:

```{r extreme}
most_seasonal <- tourism %>%
  features(Trips, feat_stl) %>%
  filter(seasonal_strength_year == max(seasonal_strength_year))
```

\pause\vspace*{-0.3cm}

```{r extreme2, fig.height=2}
tourism %>%
  right_join(most_seasonal, by = c("State", "Region", "Purpose")) %>%
  ggplot(aes(x = Quarter, y = Trips)) +
  geom_line() +
  facet_grid(vars(State, Region, Purpose))
```

## Feature extraction and statistics
\fontsize{9}{9}\sf

Find the most trended time series:

```{r extreme3}
most_trended <- tourism %>%
  features(Trips, feat_stl) %>%
  filter(trend_strength == max(trend_strength))
```

\pause\vspace*{-0.3cm}

```{r extreme4, fig.height=2}
tourism %>%
  right_join(most_trended, by = c("State", "Region", "Purpose")) %>%
  ggplot(aes(x = Quarter, y = Trips)) +
  geom_line() +
  facet_grid(vars(State, Region, Purpose))
```

```{r pca, echo=FALSE}
# Save pdf figures
savepdf <- function(file, width = 16, height = 10) {
  fname <<- paste("figs/", file, ".pdf", sep = "")
  pdf(fname, width = width / 2.54, height = height / 2.54, pointsize = 10)
  par(mgp = c(2.2, 0.45, 0), tcl = -0.4, mar = c(3.3, 3.6, 1.1, 1.1))
}
endpdf <- function() {
  crop::dev.off.crop(fname)
}
# Compute features
tourism_features <- tourism %>%
  features(Trips, feature_set(pkgs = "feasts"))
# Compute PCs
pcs <- tourism_features %>%
  select(-State, -Region, -Purpose) %>%
  prcomp(scale = TRUE) %>%
  broom::augment(tourism_features)
# Save some PC plots
savepdf("pca1", 18, 10)
pcs %>% ggplot(aes(x = .fittedPC1, y = .fittedPC2)) +
  geom_point() +
  theme(aspect.ratio = 1)
endpdf()
savepdf("pca2", 18, 10)
pcs %>% ggplot(aes(x = .fittedPC1, y = .fittedPC2, col = State)) +
  geom_point() +
  theme(aspect.ratio = 1)
endpdf()
savepdf("pca3", 18, 10)
pcs %>% ggplot(aes(x = .fittedPC1, y = .fittedPC2, col = Purpose)) +
  geom_point() +
  theme(aspect.ratio = 1)
endpdf()
# Find outliers
outliers <- pcs %>%
  filter(.fittedPC1 > 12 | (.fittedPC1 > 10 & .fittedPC2 > 0))
savepdf("pca4", 18, 10)
pcs %>% ggplot(aes(x = .fittedPC1, y = .fittedPC2, col = Purpose)) +
  geom_point() +
  theme(aspect.ratio = 1) +
  geom_point(data = outliers, aes(x = .fittedPC1, y = .fittedPC2), col = "black", shape = 1, size = 3)
endpdf()
```

## Feature extraction and statistics
\fontsize{9}{9}\sf

```{r tourismfeatures, eval=FALSE}
tourism_features <- tourism %>%
  features(Trips, feature_set(pkgs = "feasts"))
```

```{r echo=FALSE}
tourism_features
```

\begin{textblock}{5}(9.6,1.5)
\begin{alertblock}{}\fontsize{10}{12}\sf
All features from the feasts package
\end{alertblock}
\end{textblock}

## Feature extraction and statistics
\fontsize{8}{8}\sf

```{r pcatable}
pcs <- tourism_features %>%
  select(-State, -Region, -Purpose) %>%
  prcomp(scale = TRUE) %>%
  broom::augment(tourism_features)
```

```{r echo=FALSE}
pcs
```

\begin{textblock}{5}(9.6,1.5)
\begin{alertblock}{}\fontsize{10}{12}\sf
Principal components based on all features from the feasts package
\end{alertblock}
\end{textblock}

## Feature extraction and statistics
\fontsize{9}{9}\sf

\begin{textblock}{3.3}(.4,3)
\begin{alertblock}{}\fontsize{10}{12}\sf
Principal components based on all features from the feasts package
\end{alertblock}
\end{textblock}

```r
pcs %>% ggplot(aes(x=.fittedPC1, y=.fittedPC2)) +
  geom_point() + theme(aspect.ratio=1)
```

\placefig{4}{2.6}{height=6.4cm, width=12cm}{pca1}
\vspace*{10cm}

## Feature extraction and statistics
\fontsize{9}{9}\sf

\begin{textblock}{3.3}(.4,3)
\begin{alertblock}{}\fontsize{10}{12}\sf
Principal components based on all features from the feasts package
\end{alertblock}
\end{textblock}

```r
pcs %>% ggplot(aes(x=.fittedPC1, y=.fittedPC2, col=State)) +
  geom_point() + theme(aspect.ratio=1)
```

\placefig{4}{2.6}{height=6.4cm, width=12cm}{pca2}
\vspace*{10cm}

## Feature extraction and statistics
\fontsize{9}{9}\sf

\begin{textblock}{3.3}(.4,3)
\begin{alertblock}{}\fontsize{10}{12}\sf
Principal components based on all features from the feasts package
\end{alertblock}
\end{textblock}

```r
pcs %>% ggplot(aes(x=.fittedPC1, y=.fittedPC2, col=Purpose)) +
  geom_point() + theme(aspect.ratio=1)
```

\only<1>{\placefig{4}{2.6}{height=6.4cm, width=12cm}{pca3}}
\only<2>{\placefig{4}{2.6}{height=6.4cm, width=12cm}{pca4}}
\vspace*{10cm}

## Feature extraction and statistics
\fontsize{8}{8}\sf

```{r outliers2, fig.height=4.7, fig.width=12, out.width="90%"}
outliers %>%
  left_join(tourism, by = c("State", "Region", "Purpose")) %>%
  mutate(Series = glue("{State}", "{Region}", "{Purpose}", .sep = "\n\n")) %>%
  ggplot(aes(x = Quarter, y = Trips)) + geom_line() +
  facet_grid(Series ~ ., scales = "free_y") +
  ggtitle("Outlying time series in PC space")
```


# fable: Forecasting

```{r austa, include=FALSE}
# Grab ABS data
austa <- readxl::read_excel("data/340101.xls", sheet = "Data1", skip = 9) %>%
  rename(date = `Series ID`, value = A85375847A) %>%
  select(date, value) %>%
  transmute(
    Month = yearmonth(date),
    Visitors = value / 1e3
  ) %>%
  bind_rows(tibble(
    Month = yearmonth(seq(as.Date("2021-06-01"), by = "1 month", length = 7)),
    Visitors = NA_real_
  )) %>%
  as_tsibble(index = Month) %>%
  filter(Month >= yearmonth("2000 Jan"))
# Fit ETS model
fit <- austa %>%
  filter(Month < yearmonth("2018 Jan")) %>%
  model(ETS(Visitors))
# Product forecasts
fc <- forecast(fit, h = 48) %>%
  mutate(Month = as.Date(Month))
# Simulate 100 future sample paths
set.seed(1967)
sim <- fit %>%
  generate(h = 48, times = 100) %>%
  mutate(
    replicate = factor(.rep, levels = 1:100, labels = paste("Future", 1:100)),
    .rep = as.numeric(.rep)
  ) %>%
  as_tibble() %>%
  mutate(Month = as.Date(Month))
# Nice colors
cols <- scale_color_manual(
  values = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442", sample(rainbow(93))),
  breaks = paste("Future", 1:100),
  name = " "
)
# Now build up some plots with alignment
p1 <- austa %>%
  mutate(Visitors = if_else(Month >= yearmonth("2018 Jan"), NA_real_, Visitors)) %>%
  as_tibble() %>%
  mutate(Month = as.Date(Month)) %>%
  ggplot(aes(x = Month, y = Visitors)) +
  geom_line() +
  labs(
    x = "Month",
    y = "Thousands of visitors",
    title = "Total short-term visitors to Australia"
  ) +
  scale_x_date(
    breaks = seq(as.Date("2000-01-01"), by = "5 years", l = 5),
    labels = paste("Jan", seq(2000, 2020, by = 5)),
    minor_breaks = seq(as.Date("2001-01-01"), by = "1 year", l = 25)
  ) +
  ylim(min(austa$Visitors, sim$.sim, na.rm = TRUE), max(austa$Visitors, sim$.sim, na.rm = TRUE))
p2 <- p1 + cols +
  geom_line(aes(y = .sim, group = replicate, color = replicate),
    data = sim %>% filter(.rep <= 1)
  )
p3 <- p1 + cols +
  geom_line(aes(y = .sim, group = replicate, color = replicate),
    data = sim %>% filter(.rep <= 2)
  )
p4 <- p1 + cols +
  geom_line(aes(y = .sim, group = replicate, color = replicate),
    data = sim %>% filter(.rep <= 3)
  )
p5 <- p1 + cols +
  geom_line(aes(y = .sim, group = replicate, color = replicate),
    data = sim %>% filter(.rep <= 10)
  )
p6 <- p1 + cols +
  geom_line(aes(y = .sim, group = replicate, color = replicate), alpha = 0.5, data = sim) +
  guides(color = "none")
p7 <- p1 +
  geom_line(aes(y = .sim, group = replicate, color = replicate), alpha = 0.5, col = "gray", data = sim) +
  guides(color = "none")
p8 <- p7 + autolayer(fc, level = c(50, 90))
p9 <- p8 + coord_cartesian(xlim = as.Date(c("2013-01-01", "2021-08-01"))) +
  scale_x_date(
    breaks = seq(as.Date("2000-01-01"), by = "2 years", l = 11),
    labels = paste("Jan", seq(2000, 2020, by = 2)),
    minor_breaks = seq(as.Date("2001-01-01"), by = "1 year", l = 25)
  )
p10 <- p9 +
  geom_line(
    color = "black",
    data = austa %>% filter(Month >= yearmonth("2018 Jan")) %>% mutate(Month = as.Date(Month))
  )
aligned_plots <- align_patches(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10)

for (i in seq(10)) {
  pdf(paste0("figs/austa", i, ".pdf"), height = 3.3, width = 8)
  print(aligned_plots[[i]])
  dev.off()
}
```

## Random futures

\forecast\pause
\includegraphics[height=8cm]{austa1}

## Random futures

\forecast
\includegraphics[height=8cm]{austa2}
\simfutures

## Random futures

\forecast
\includegraphics[height=8cm]{austa3}
\simfutures

## Random futures

\forecast
\includegraphics[height=8cm]{austa4}
\simfutures

## Random futures

\forecast
\includegraphics[height=8cm]{austa5}
\simfutures

## Random futures

\forecast
\includegraphics[height=8cm]{austa6}
\simfutures

## Random futures

\forecast
\includegraphics[height=8cm]{austa7}
\simfutures

## Random futures

\forecast
\includegraphics[height=8cm]{austa8}
\simfutures

## Random futures

\forecast
\includegraphics[height=8cm]{austa9}
\simfutures

## Random futures

\forecast
\includegraphics[height=8cm]{austa10}
\simfutures

## Model fitting
\fontsize{10}{11}\sf

```{r holiday-model}
holiday_fit <- holidays %>%
  model(
    snaive = SNAIVE(Trips),
    naive = NAIVE(Trips),
    ets = ETS(Trips),
    arima = ARIMA(Trips)
  )
```


```{r holiday-model2, echo=FALSE, dependson='holiday-model'}
holiday_fit
```

## Model fitting
\fontsize{10}{11}\sf

```{r holiday-model3, dependson='holiday-model'}
holiday_fit %>%
  filter(State == "VIC") %>%
  select(arima) %>%
  report()
```

## Model fitting
\fontsize{10}{11}\sf

```{r holiday-model4, dependson='holiday-model'}
glance(holiday_fit)
```

## Model fitting
\fontsize{10}{11}\sf

```{r holiday-model5, dependson='holiday-model'}
tidy(holiday_fit)
```

## Model fitting
\fontsize{10}{11}\sf

```{r holiday-model6, dependson='holiday-model'}
augment(holiday_fit)
```

## Ljung-Box test

```{r lbtest, echo=TRUE, fig.height=4, warning = FALSE}
augment(holiday_fit) %>%
  filter(State == "VIC", .model == "arima") %>%
  features(.resid, ljung_box, dof = 2, lag = 8)
```

## `gg_tsresiduals()` function
\fontsize{9}{10}\sf

```{r ggtsresiduals, echo=TRUE, warning = FALSE, fig.height=3}
holiday_fit %>%
  filter(State == "VIC") %>%
  select(arima) %>%
  gg_tsresiduals()
```

## Producing forecasts

\fontsize{9}{9}\sf

```{r holiday-fc, echo = TRUE, dependson='holiday-model'}
holiday_fc <- holiday_fit %>%
  forecast(h = "2 years")
```

```{r holiday-fbl, echo = FALSE, dependson='holiday-fc'}
print(holiday_fc, n = 10)
```

## Visualising forecasts
\fontsize{9}{9}\sf

```{r holiday-fc-plot, warning=FALSE, message=FALSE, fig.height=2.9, dependson='holiday-fc'}
holiday_fc %>%
  filter(State == "VIC") %>%
  autoplot(holidays, level = NULL) +
  labs(title = "Holidays in Victoria", y = "Thousands of visitors") +
  guides(color = guide_legend(title = "Forecast"))
```


## Visualising forecasts
\fontsize{9}{9}\sf

```{r holiday-fc-plot2, warning=FALSE, message=FALSE, fig.height=2.9, dependson='holiday-fc'}
holiday_fc %>%
  filter(State == "VIC", .model == "arima") %>%
  autoplot(holidays) +
  labs(title = "Holidays in Victoria", y = "Thousands of visitors") +
  guides(color = guide_legend(title = "Forecast"))
```

## Prediction intervals
\fontsize{10}{12}\sf

```{r holiday-fc-interval, dependson='holiday-fc'}
holiday_fc %>% hilo(level = 95)
```

## Prediction intervals

 * Point forecasts often useless without a measure of uncertainty (such as prediction intervals).
 * Prediction intervals require a stochastic model (with random errors, etc).
 * For most models, prediction intervals get wider as the forecast horizon increases.
 * Use `level` argument to control coverage.
 * Check residual assumptions before believing them.
 * Usually too narrow due to unaccounted uncertainty.

# fable: Evaluating forecast accuracy

## Training and test sets

```{r traintest, fig.height=1, echo=FALSE, cache=TRUE}
train <- 1:18
test <- 19:24
par(mar = c(0, 0, 0, 0))
plot(0, 0, xlim = c(0, 26), ylim = c(0, 2), xaxt = "n", yaxt = "n", bty = "n", xlab = "", ylab = "", type = "n")
arrows(0, 0.5, 25, 0.5, 0.05)
points(train, train * 0 + 0.5, pch = 19, col = "#0072B2")
points(test, test * 0 + 0.5, pch = 19, col = "#D55E00")
text(26, 0.5, "time")
text(10, 1, "Training data", col = "#0072B2")
text(21, 1, "Test data", col = "#D55E00")
```

-   A model which fits the training data well will not necessarily forecast well.
-   A perfect fit can always be obtained by using a model with enough parameters.
-   Over-fitting a model to data is just as bad as failing to identify a systematic pattern in the data.
  * The test set must not be used for *any* aspect of model development or calculation of forecasts.
  * Forecast accuracy is based only on the test set.

## Measures of forecast accuracy

```{r holidays-fc-1, echo=FALSE, fig.height=4}
vic_holidays <- holidays %>% filter(State == "VIC")
train <- vic_holidays %>% filter(year(Quarter) <= 2015)
vic_fit <- train %>%
  model(
    snaive = SNAIVE(Trips),
    naive = NAIVE(Trips),
    ets = ETS(Trips),
    arima = ARIMA(Trips)
  )
vic_fc <- vic_fit %>%
  forecast(h = 8)
vic_fc %>%
  autoplot(vic_holidays, level = NULL) +
  labs(title = "Holidays in Victoria", y = "Thousands of visitors") +
  guides(color = guide_legend(title = "Forecast"))
```

## Measures of forecast accuracy
\fontsize{10}{10}\sf

```{r holidays-train-accuracy}
accuracy(vic_fc, holidays)
```

```{r tscvplots, echo=FALSE}
tscv_plot <- function(.init, .step, h = 1) {
  expand.grid(
    time = seq(26),
    .id = seq(trunc(20 / .step))
  ) %>%
    group_by(.id) %>%
    mutate(
      observation = case_when(
        time <= ((.id - 1) * .step + .init) ~ "train",
        time %in% c((.id - 1) * .step + .init + h) ~ "test",
        TRUE ~ "unused"
      )
    ) %>%
    ungroup() %>%
    filter(.id <= 26 - .init) %>%
    ggplot(aes(x = time, y = .id)) +
    geom_segment(
      aes(x = 0, xend = 27, y = .id, yend = .id),
      arrow = arrow(length = unit(0.015, "npc")),
      col = "black", size = .25
    ) +
    geom_point(aes(col = observation), size = 2) +
    scale_y_reverse() +
    scale_color_manual(values = c(train = "#0072B2", test = "#D55E00", unused = "gray")) +
    # theme_void() +
    # geom_label(aes(x = 28.5, y = 1, label = "time")) +
    guides(col = FALSE) +
    labs(x = "time", y = "") +
    theme_void() +
    theme(axis.title = element_text())
}
```

## Time series cross-validation {-}

**Traditional evaluation**

```{r traintest1, fig.height=1, echo=FALSE, dependson="tscvplots", message=FALSE}
tscv_plot(.init = 18, .step = 20, h = 1:8) +
  geom_text(aes(x = 10, y = 0.8, label = "Training data"), color = "#0072B2") +
  geom_text(aes(x = 21, y = 0.8, label = "Test data"), color = "#D55E00") +
  ylim(1, 0)
```

\pause

**Time series cross-validation**

```{r tscvggplot1, echo=FALSE, message=FALSE}
tscv_plot(.init = 3, .step = 1, h = 1) +
  geom_text(aes(x = 21, y = 0, label = "h = 1"), color = "#D55E00")
```


## Time series cross-validation {-}

**Traditional evaluation**

```{r traintest2, ref.label="traintest1", fig.height=1, echo=FALSE, message=FALSE}
```

**Time series cross-validation**

```{r tscvggplot2, echo=FALSE,  dependson="tscvplots", message=FALSE}
tscv_plot(.init = 3, .step = 1, h = 2) +
  geom_text(aes(x = 21, y = 0, label = "h = 2"), color = "#D55E00")
```

## Time series cross-validation {-}

**Traditional evaluation**

```{r traintest3, ref.label="traintest1", fig.height=1, echo=FALSE, message=FALSE}
```

**Time series cross-validation**

```{r tscvggplot3, echo=FALSE,  dependson="tscvplots", message=FALSE}
tscv_plot(.init = 3, .step = 1, h = 3) +
  geom_text(aes(x = 21, y = 0, label = "h = 3"), color = "#D55E00")
```


## Time series cross-validation {-}

**Traditional evaluation**

```{r traintest4, ref.label="traintest1", fig.height=1, echo=FALSE, message=FALSE}
```

**Time series cross-validation**

```{r tscvggplot4, echo=FALSE,  dependson="tscvplots", message=FALSE}
tscv_plot(.init = 3, .step = 1, h = 4) +
  geom_text(aes(x = 21, y = 0, label = "h = 4"), color = "#D55E00")
```

\only<2>{\begin{textblock}{6.7}(.3,6.3)\begin{block}{}\fontsize{12}{13}\sf
\begin{itemize}\tightlist
\item Forecast accuracy averaged over test sets.
\item Also known as "evaluation on a rolling forecasting origin"
\end{itemize}\end{block}\end{textblock}}

\vspace*{10cm}

## Time series cross-validation {-}

\fontsize{12}{13}\sf

Stretch with a minimum length of 4 years, growing by 1 quarter each step.

```{r vichol-stretch}
vic_holiday_stretch <- holidays %>%
  filter(State == "VIC") %>%
  stretch_tsibble(.init = 16, .step = 1)
```
\fontsize{10}{11}\sf
```{r vichol-stretch-print, echo = FALSE, dependson="vichol-stretch"}
vic_holiday_stretch
```

## Time series cross-validation {-}

\fontsize{10}{11}\sf

```{r vichol-fit, dependson="vichol-stretch"}
fit_cv <- vic_holiday_stretch %>%
  model(
    ets = ETS(Trips),
    arima = ARIMA(Trips),
    snaive = SNAIVE(Trips)
  )
```

\fontsize{10}{11}\sf
```{r vichol-fit-print, echo = FALSE, dependson="vichol-fit"}
print(fit_cv, n = 6)
```

## Time series cross-validation {-}

\fontsize{10}{11}\sf

Produce one step ahead forecasts from all models.

```{r vichol-fc, dependson="vichol-fit"}
fc_cv <- fit_cv %>%
  forecast(h = 1)
```

```{r vichol-fc-print, echo = FALSE, dependson="vichol-fc"}
fc_cv %>%
  select(-.model) %>%
  print(n = 6)
```

## Time series cross-validation {-}

\fontsize{10}{11}\sf

```{r vichol-accuracy, dependson="vichol-fc"}
# Cross-validated
fc_cv %>% accuracy(holidays)
```

\fontsize{11}{11}\sf

```{r, echo = FALSE, warning = FALSE, eval=FALSE, dependson="vichol-fc"}
fc_cv %>%
  accuracy(holidays) %>%
  mutate(.type = "Cross-validation") %>%
  bind_rows(
    fb_stock %>%
      model(RW(Close ~ drift())) %>%
      accuracy()
  ) %>%
  transmute(Type = .type, RMSE, MAE, MAPE) %>%
  gt::gt("Type") %>%
  gt::fmt_number(columns = 2:4, decimals = 3) %>%
  gt::as_latex()
```

A good way to choose the best forecasting model is to find the model with the smallest RMSE computed using time series cross-validation.

# fable: Forecast reconciliation

## Hierarchical time series
\fontsize{13}{14}\sf

A \alert{\textbf{hierarchical time series}} is a collection of several time series that are linked together in a hierarchical structure.

\begin{minipage}{9.6cm}
\begin{block}{}
\begin{tikzpicture}
\tikzstyle{every node}=[ellipse,draw,inner sep=0.2pt,fill=red!15]
\tikzstyle[level distance=.1cm]
\tikzstyle[sibling distance=7cm]
\tikzstyle{level 1}=[sibling distance=33mm,set style={{every node}+=[fill=blue!15]}]
\tikzstyle{level 2}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=yellow]}]
\node{Total}[edge from parent fork down]
 child {node {A}
   child {node {AA}}
   child {node {AB}}
   child {node {AC}}
 }
 child {node {B}
   child {node {BA}}
   child {node {BB}}
   child {node {BC}}
 }
 child {node {C}
   child {node {CA}}
   child {node {CB}}
   child {node {CC}}
 };
\end{tikzpicture}
\end{block}
\end{minipage}

\pause\alert{Examples}\vspace*{-0.2cm}

 * Tourism demand by states, zones, regions

## Grouped time series
\fontsize{13}{14}\sf

A \alert{\textbf{grouped time series}} is a collection of time series that can be grouped together in a number of non-hierarchical ways.

\begin{minipage}{9.2cm}
\begin{block}{}
\begin{tikzpicture}[level distance=1.5cm]
\tikzstyle{every node}=[ellipse,draw,inner sep=0.2pt,outer sep=0pt, fill=red!15]
\tikzstyle{level 1}=[sibling distance=23mm,set style={{every node}+=[fill=blue!15]},level distance=1cm]
\tikzstyle{level 2}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=yellow]}, level distance=0.9cm]
\node{Total}[edge from parent fork down]
 child {node {A}
   child {node {AX}}
   child {node {AY}}
 }
 child {node {B}
   child {node {BX}}
   child {node {BY}}
 };
\end{tikzpicture}\hspace*{1cm}
\begin{tikzpicture}[level distance=1.5cm]
\tikzstyle{every node}=[ellipse,draw,inner sep=0.2pt,outer sep=0pt, fill=red!15]
\tikzstyle{level 1}=[sibling distance=23mm,set style={{every node}+=[fill=blue!15]},level distance=1cm]
\tikzstyle{level 2}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=yellow]}, level distance=0.9cm]
\node{Total}[edge from parent fork down]
 child {node {X}
   child {node {AX}}
   child {node {BX}}
 }
 child {node {Y}
   child {node {AY}}
   child {node {BY}}
 };
\end{tikzpicture}
\end{block}
\end{minipage}

\pause\alert{Examples}

 * Tourism by state and purpose of travel
 * Retail sales by product groups/sub groups, and by countries/regions

## Creating aggregates
\fontsize{8}{8}\sf

```{r tourism_aggregate}
tourism %>%
  aggregate_key(Purpose * (State / Region), Trips = sum(Trips)) %>%
  filter(Quarter == yearquarter("1998 Q1")) %>%
  print(n = 15)
```

## Creating aggregates
\fontsize{13}{15}\sf

 * Similar to `summarise()` but using the key structure
 * A grouped structure is specified using `grp1 * grp2`
 * A nested structure is specified via `parent / child`.
 * Groups and nesting can be mixed:

    ```r
    (country/region/city) * (brand/product)
    ```

 * All possible aggregates are produced.
 * These are useful when forecasting at different levels of aggregation.


## The problem
\fontsize{13}{14}\sf

\begin{alertblock}{}
\begin{enumerate}\tightlist
 \item How to forecast time series at all nodes such that the forecasts add up in the same way as the original data?
 \item Can we exploit relationships between the series to improve the forecasts?
\end{enumerate}
\end{alertblock}\pause

### The solution

1. Forecast all series at all levels of aggregation using an automatic forecasting algorithm.\newline (e.g., `ETS`, `ARIMA`, ...)
2. Reconcile the resulting forecasts so they add up correctly using least squares optimization (i.e., find closest reconciled forecasts to the original forecasts).
3. This is available using `reconcile()`.

## Forecast reconciliation
\fontsize{9}{10}\sf

```{r tourismets_reconciled, message=FALSE}
tourism %>%
  aggregate_key(Purpose * (State / Region), Trips = sum(Trips)) %>%
  model(ets = ETS(Trips)) %>%
  reconcile(ets_adjusted = min_trace(ets)) %>%
  forecast(h = 2)
```

## Hierarchical and grouped time series

Every collection of time series with aggregation constraints can be written as
\begin{block}{}
\centerline{$\by_{t}=\bS\bm{b}_{t}$}
\end{block}
where

 * $\by_t$ is a vector of all series at time $t$
 * $\bm{b}_t$ is a vector of the most disaggregated series at time $t$
 * $\bS$ is a ``summing matrix'' containing the aggregation constraints.

## Hierarchical time series

\begin{minipage}{4cm}\vspace*{0.2cm}
\begin{block}{}\centering
\begin{tikzpicture}
\tikzstyle{every node}=[ellipse,draw,fill=red!15,inner sep=2pt]
\tikzstyle[level distance=.3cm]
\tikzstyle[sibling distance=12cm]
\tikzstyle{level 1}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=blue!15]}]
\node{Total}[edge from parent fork down]
 child {node {A}
 }
 child {node {B}
 }
 child {node {C}
 };
\end{tikzpicture}
\end{block}
\end{minipage}

\only<2->{\begin{textblock}{6.3}(6,1)\small
\begin{itemize}\itemsep=0cm\parskip=0cm
\item[$ y_{t}: $] observed aggregate of all series at time
$t$.
\item[$ y_{X,t}: $] observation on series $X$ at time $t$.
\item[$ \bm{b}_{t}: $] vector of all series at bottom level
in time $t$.
\end{itemize}
\end{textblock}}\vspace*{0.6cm}
\only<3->{
$\bY_{t}= \begin{pmatrix}
  y_{t}\\
  y_{A,t}\\
  y_{B,t}\\
  y_{C,t}
  \end{pmatrix} = \only<3>{\hspace*{0.01cm}\begin{pmatrix}
                1 & 1 & 1 \\
                1 & 0 & 0 \\
                0 & 1 & 0\\
                0 & 0 & 1
                \end{pmatrix}}\only<4->{{\color{blue}\underbrace{\begin{pmatrix}
                1 & 1 & 1 \\
                1 & 0 & 0 \\
                0 & 1 & 0\\
                0 & 0 & 1
                \end{pmatrix}}_{\bS}}}\only<3>{\hspace*{0.08cm}}\only<3>{\hspace*{-0.1cm}\begin{pmatrix}y_{A,t}\\y_{B,t}\\y_{C,t}\end{pmatrix}}\rule{0cm}{1.6cm}
                \only<4->{\hspace*{0.08cm}{\color{red}\underbrace{\begin{pmatrix}y_{A,t}\\y_{B,t}\\y_{C,t}\end{pmatrix}}_{\bm{b}_{t}}}}$}

\vspace*{-0.6cm}

\only<4>{\hspace*{8cm}\colorbox[RGB]{210,210,210}{$\bY_{t}=\color{blue}\bS\color{red}\bm{b}_{t}$}}

\vspace*{10cm}

## Forecasting notation

Let $\hat{\by}_n(h)$ be vector of initial $h$-step forecasts, made at time $n$, stacked in same order as $\by_t$. \pause\newline  (In general, they will not ``add up''.)\pause

\begin{block}{}
Reconciled forecasts must be of the form:
$$\tilde{\by}_{n}(h)=\bS\bm{G}\hat{\by}_{n}(h)$$
for some matrix $\bm{G}$.
\end{block}\pause

 * $\bm{G}$ extracts and combines base forecasts $\hat{\by}_{n}(h)$ to get bottom-level forecasts.
 * $\bS$ adds them up

## Optimal combination forecasts
\fontsize{13}{14}\sf

\begin{alertblock}{Main result}
The best (minimum sum of variances) unbiased forecasts are obtained when
$\bm{G} = (\bS'\bW^{-1}_{h}\bS)^{-1}\bS'\bW^{-1}_{h}$,
where $\bW_h$ is the $h$-step base forecast error covariance matrix.
\end{alertblock}

\pause

\begin{block}{}
\centerline{$\displaystyle\textcolor{red}{\tilde{\by}_{n}(h)}
=\bS(\bS'\bW^{-1}_{h}\bS)^{-1}\bS'\bW^{-1}_{h}\textcolor{blue}{\hat{\by}_{n}(h)}$}
\end{block}\fontsize{14}{15}\sf

\alert{\textbf{Problem:}} $\bW_h$ hard to estimate, especially for $h>1$.

\alert{Solutions:}\vspace*{-0.4cm}

 * Ignore $\bW_h$ (OLS) [`min_trace(method='ols')`]
 * Assume $\bW_h = k_h\bW_1$ is diagonal (WLS) [`min_trace(method='wls')`]
 * Assume $\bW_h = k_h\bW_1$ and estimate it (GLS) [`min_trace(method='shrink')` (the default)]

## Features
\fontsize{15}{17}\sf

 * Covariates can be included in initial forecasts.
 * Adjustments can be made to initial forecasts at any level.
 * Very simple and flexible method. Can work with *any* hierarchical or grouped time series.
 * Conceptually easy to implement: regression of base forecasts on structure matrix.

## Example: Australian tourism
\fontsize{12}{13}\sf

```{r fctourism}
tourism_agg <- tourism %>%
  aggregate_key(Purpose * (State / Region),
    Trips = sum(Trips)
  )
fc <- tourism_agg %>%
  filter_index(. ~ "2015 Q4") %>%
  model(ets = ETS(Trips)) %>%
  reconcile(ets_adjusted = min_trace(ets)) %>%
  forecast(h = "2 years")
```

## Example: Australian tourism
\fontsize{10}{11}\sf

```{r fctourism2, dependson='fctourism'}
fc %>%
  filter(is_aggregated(Purpose) & is_aggregated(State)) %>%
  autoplot(tourism_agg, level = 95)
```

## Example: Australian tourism
\fontsize{10}{11}\sf

```{r fctourism3, dependson='fctourism'}
fc %>%
  filter(is_aggregated(Purpose) & State == "VIC" &
    is_aggregated(Region)) %>%
  autoplot(tourism_agg, level = 95)
```

## Example: Australian tourism
\fontsize{10}{11}\sf

```{r fctourism4, dependson='fctourism'}
fc %>%
  filter(is_aggregated(Purpose) & Region == "Melbourne") %>%
  autoplot(tourism_agg, level = 95)
```

## Example: Australian tourism
\fontsize{10}{11}\sf

```{r fctourism5, dependson='fctourism'}
fc %>%
  filter(is_aggregated(Purpose) & Region == "Snowy Mountains") %>%
  autoplot(tourism_agg, level = 95)
```

## Example: Australian tourism
\fontsize{10}{11}\sf

```{r fctourism6, dependson='fctourism'}
fc %>%
  filter(Purpose == "Holiday" & Region == "Barossa") %>%
  autoplot(tourism_agg, level = 95)
```

## Example: Australian tourism
\fontsize{10}{11}\sf

```{r fctourism7, dependson='fctourism'}
fc %>%
  filter(is_aggregated(Purpose) & Region == "MacDonnell") %>%
  autoplot(tourism_agg, level = 95)
```

## Example: Australian tourism
\fontsize{10}{10}\sf

```{r fctourismcomb}
fc <- tourism_agg %>%
  filter_index(. ~ "2015 Q4") %>%
  model(
    ets = ETS(Trips),
    arima = ARIMA(Trips)
  ) %>%
  mutate(
    comb = (ets + arima) / 2
  ) %>%
  reconcile(
    ets_adj = min_trace(ets),
    arima_adj = min_trace(arima),
    comb_adj = min_trace(comb)
  ) %>%
  forecast(h = "2 years")
```

## Forecast evaluation
\fontsize{10}{11}\sf

```{r fcaccuracy, dependson='fctourismcomb'}
fc %>% accuracy(tourism_agg)
```

## Forecast evaluation
\fontsize{11}{11}\sf

```{r fcaccuracy2, dependson='fctourismcomb'}
fc %>%
  accuracy(tourism_agg) %>%
  group_by(.model) %>%
  summarise(MASE = mean(MASE)) %>%
  arrange(MASE)
```

## Tidyverts developers

\begin{textblock}{5.7}(0.4,1.4)
\begin{block}{Earo Wang}
\includegraphics[height=5.8cm]{earowang}
\end{block}
\end{textblock}

\begin{textblock}{5.7}(6.75,1.4)
\begin{block}{Mitchell O'Hara-Wild}
\includegraphics[height=5.8cm]{mitch2}
\end{block}
\end{textblock}

## More information
\fontsize{18}{20}\sf

 * Slides and papers: **robjhyndman.com**
 * Packages: **tidyverts.org**
 * Forecasting textbook using fable package: **OTexts.com/fpp3**

\begin{textblock}{8}(7.6,4.8)
\begin{alertblock}{Find me at \dots}
\href{https://twitter.com/robjhyndman}{\faicon{twitter} @robjhyndman}

\href{https://github.com/robjhyndman}{\faicon{github}  @robjhyndman}

\href{https://robjhyndman.com}{\faicon{home} robjhyndman.com}

\href{mailto:rob.hyndman@monash.edu}{\faicon{envelope}  rob.hyndman@monash.edu}
\end{alertblock}
\end{textblock}
\vspace*{10cm}
