---
title: "Feasts & fables: time series analysis using R"
author: "Rob J Hyndman"
date: "RBA 13 March 2023"
toc: true
classoption: aspectratio=169
output:
  binb::monash:
    colortheme: monashwhite
    fig_width: 8
    fig_height: 3
    keep_tex: no
    includes:
      in_header: header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  dev.args = list(pointsize = 11)
)
# Load packages
library(fpp3)
library(patchwork)
library(glue)
library(stringr)
scale_colour_discrete <- function(...) {
  scale_colour_manual(..., values = c("#0072b2", "#cc79a7", "#009e73","#e69f00",  "#f0e442", "#56b4e9", "#d55e00", "#666666"))
}
scale_fill_discrete <- function(...) {
  scale_fill_manual(..., values = c("#0072b2", "#cc79a7", "#009e73","#e69f00",  "#f0e442", "#56b4e9", "#d55e00", "#666666"))
}

options(
  digits = 3,
  width = 80
)

# Neater data sets
tourism <- tsibble::tourism |>
  mutate(
    State = recode(State,
      "Australian Capital Territory" = "ACT",
      "New South Wales" = "NSW",
      "Northern Territory" = "NT",
      "Queensland" = "QLD",
      "South Australia" = "SA",
      "Tasmania" = "TAS",
      "Victoria" = "VIC",
      "Western Australia" = "WA"
    ),
    State = factor(State, levels=c("NSW","VIC","QLD","WA","SA","TAS","NT","ACT")),
    Region = str_remove(Region, "Australia's ")
  )
```

# tsibble: Time series data

## Tidyverts packages

\begin{textblock}{3.8}(11,0)\begin{alertblock}{}\Large\textbf{tidyverts.org}\end{alertblock}\end{textblock}

\placefig{10}{1.4}{width=2.5cm}{tsibble.png}
\placefig{10}{4.4}{width=2.5cm}{feasts.png}
\placefig{10}{7.4}{width=2.5cm}{fable.png}

\begin{textblock}{8}(0.4,1.2)
\fontsize{8}{9}\sf

```r
library(fpp3)
```

\vspace*{-0.4cm}
\begin{alltt}
\input{startuppdf.txt}
\end{alltt}
\end{textblock}

## Australian tourism regions

```{r ausmap, echo=FALSE, message=FALSE, width=20, height=18, out.height="90%"}
library(sf)
# Use Okabe-Ito color-blind friendly color palette
state_colors <- c(
  `Australian Capital Territory` = "#666666",
  `New South Wales` = "#0072b2",
  `Northern Territory` = "#d55e00",
  `Queensland` = "#009e73",
  `South Australia` = "#f0e442",
  `Tasmania` = "#56b4e9",
  `Victoria` = "#cc79a7",
  `Western Australia` = "#e69f00"
)
read_sf("tourism/Tourism_Regions_2020.shp") |>
  rename(State = "STE_NAME16") |>
  mutate(State = factor(State, levels = c(
    "Queensland","Northern Territory","Western Australia",
    "South Australia","New South Wales","Australian Capital Territory",
    "Victoria","Tasmania"
  ))) |> 
  ggplot() +
  geom_sf(aes(fill = State), alpha = 0.8) +
  theme_void() +
  scale_fill_manual(values = state_colors)
```

\only<2>{\begin{textblock}{6.5}(9.1,1.4)
\begin{block}{}\fontsize{12}{12}\sf
  \begin{itemize}\itemsep=0cm\parskip=0cm
    \item Quarterly data on visitor nights: 1998 -- 2017
    \item From \textit{National Visitor Survey}, interviews of 120,000 Australians aged 15+.
    \item Geographical hierarchy split by
    \begin{itemize}
    \item 8 states and territories
    \item 76 regions
    \end{itemize}
    \item Purpose:
    \begin{itemize}
    \item Holidays
    \item Business
    \item Visiting friends \& relatives
    \item Other
    \end{itemize}
  \end{itemize}
\end{block}
\end{textblock}}

## `tsibble` objects

\fontsize{10}{11.3}\sf

```{r, echo = TRUE}
tourism
```

\only<2->{\begin{textblock}{1.1}(2.1,3.7)
\begin{alertblock}{}\fontsize{10}{10}\sf Index\phantom{dg}\end{alertblock}
\end{textblock}}
\only<3->{\begin{textblock}{3.9}(3.65,3.7)
\begin{alertblock}{}\fontsize{10}{10}\sf Keys\phantom{dg}\end{alertblock}
\end{textblock}}
\only<4>{\begin{textblock}{1.5}(7.95,3.7)
\begin{alertblock}{}\fontsize{10}{10}\sf Measure\phantom{dg}\end{alertblock}
\end{textblock}}

\begin{textblock}{3}(10,5)
\begin{block}{}\fontsize{10}{10}\sf Domestic visitor nights in thousands by state/region and purpose.\phantom{dg}\end{block}
\end{textblock}

## `tsibble` objects

* A `tsibble` allows storage and manipulation of multiple time series in R.

* It contains:

  + An index: time information about the observation
  + Measured variable(s): numbers of interest
  + Key variable(s): optional unique identifiers for each series

* It works with tidyverse functions.

# feasts: Data visualization

## Australian holidays
\fontsize{9}{10}\sf

```{r holidays}
holidays <- tourism |>
  filter(Purpose == "Holiday") |>
  group_by(State) |>
  summarise(Trips = sum(Trips))
```

```{r, echo=FALSE}
holidays
```

## Australian holidays
\fontsize{9}{10}\sf

```{r holidays-plot, echo=TRUE, dependson="holidays"}
holidays |> autoplot(Trips) +
  labs(x = "thousands of trips", title = "Australian domestic holiday nights")
```

## Seasonal plots
\fontsize{9}{10}\sf

```{r graphics1, fig.width=6, fig.height=6, out.width="40%"}
holidays |> gg_season(Trips) +
  labs(x = "thousands of trips", title = "Australian domestic holiday nights")
```

## Seasonal subseries plots
\fontsize{9}{10}\sf

```{r graphics2, fig.width=9, fig.height=5, out.height="70%"}
holidays |> gg_subseries(Trips) +
  labs(x = "thousands of trips", title = "Australian domestic holiday nights")
```

# feasts: Time series features

## Strength of seasonality and trend

\begin{alertblock}{STL decomposition}
\centerline{$y_t = T_t+S_t+R_t$}
\end{alertblock}

### Seasonal strength
$$\max\left(0, 1-\frac{\text{Var}(R_t)}{\text{Var}(S_t+R_t)}\right)$$

### Trend strength
$$\max\left(0, 1-\frac{\text{Var}(R_t)}{\text{Var}(T_t+R_t)}\right)$$

## Feature extraction and statistics

\fontsize{9}{9}\sf

```{r features}
tourism |> features(Trips, feat_stl)
```

## Feature extraction and statistics
\fontsize{8}{9}\sf

```{r features-plot, fig.height=4, fig.width=6, out.height="60%"}
tourism |>
  features(Trips, feat_stl) |>
  ggplot(aes(x = trend_strength, y = seasonal_strength_year, col = Purpose)) +
  geom_point() + 
  facet_wrap(vars(State))
```

\only<2->{\begin{textblock}{5.7}(8.8,6.8)
\begin{alertblock}{}\fontsize{12}{12}\sf
\begin{itemize}\tightlist
\item Holidays more seasonal than other travel.
\item WA has strongest trends.
\end{itemize}
\end{alertblock}\end{textblock}}

## Feature extraction and statistics
\fontsize{9}{9}\sf

Find the most seasonal time series:

```{r extreme}
most_seasonal <- tourism |>
  features(Trips, feat_stl) |>
  filter(seasonal_strength_year == max(seasonal_strength_year))
```

\pause\vspace*{-0.3cm}

```{r extreme2, fig.height=2.03}
tourism |>
  right_join(most_seasonal, by = c("State", "Region", "Purpose")) |>
  ggplot(aes(x = Quarter, y = Trips)) +
  geom_line() +
  facet_grid(vars(State, Region, Purpose))
```


```{r pca, echo=FALSE}
# Save pdf figures
savepdf <- function(file, width = 16, height = 10) {
  fname <<- paste("figs/", file, ".pdf", sep = "")
  pdf(fname, width = width / 2.54, height = height / 2.54, pointsize = 10)
  par(mgp = c(2.2, 0.45, 0), tcl = -0.4, mar = c(3.3, 3.6, 1.1, 1.1))
}
endpdf <- function() {
  crop::dev.off.crop(fname)
}
# Compute features
tourism_features <- tourism |>
  features(Trips, feature_set(pkgs = "feasts"))
# Compute PCs
pcs <- tourism_features |>
  select(-State, -Region, -Purpose) |>
  prcomp(scale = TRUE) |>
  broom::augment(tourism_features)
# Save some PC plots
savepdf("pca1", 18, 10)
pcs |> ggplot(aes(x = .fittedPC1, y = .fittedPC2)) +
  geom_point() +
  theme(aspect.ratio = 1)
endpdf()
savepdf("pca2", 18, 10)
pcs |> ggplot(aes(x = .fittedPC1, y = .fittedPC2, col = State)) +
  geom_point() +
  theme(aspect.ratio = 1)
endpdf()
savepdf("pca3", 18, 10)
pcs |> ggplot(aes(x = .fittedPC1, y = .fittedPC2, col = Purpose)) +
  geom_point() +
  theme(aspect.ratio = 1)
endpdf()
# Find outliers
outliers <- pcs |>
  filter(.fittedPC1 > 10 & .fittedPC2 > 0 & .fittedPC2 < 2) |>
  mutate(label = glue::glue("{State}/{Region}"))
savepdf("pca4", 18, 10)
pcs |> ggplot(aes(x = .fittedPC1, y = .fittedPC2, col = Purpose)) +
  geom_point() +
  theme(aspect.ratio = 1) +
  geom_point(data = outliers, aes(x = .fittedPC1, y = .fittedPC2), col = "black", shape = 1, size = 3)
endpdf()
savepdf("pca5", 18, 10)
pcs |> 
  ggplot(aes(x = .fittedPC1, y = .fittedPC2, col = Purpose)) +
  geom_point() +
  theme(aspect.ratio = 1) +
  geom_point(data = outliers, aes(x = .fittedPC1, y = .fittedPC2), col = "black", shape = 1, size = 3) +
  ggrepel::geom_label_repel(aes(label=label), data = outliers, show.legend = FALSE)
endpdf()
```

## Feature extraction and statistics
\fontsize{9}{8}\sf

```{r tourismfeatures, eval=FALSE}
tourism_features <- tourism |>
  features(Trips, feature_set(pkgs = "feasts"))
```

```{r echo=FALSE}
tourism_features
```

\begin{textblock}{5}(9.6,1.5)
\begin{alertblock}{}\fontsize{10}{12}\sf
All features from the feasts package
\end{alertblock}
\end{textblock}

## Feature extraction and statistics
\fontsize{9}{8}\sf

```{r pcatable}
pcs <- tourism_features |>
  select(-State, -Region, -Purpose) |>
  prcomp(scale = TRUE) |>
  broom::augment(tourism_features)
```

```{r echo=FALSE}
pcs
```

\begin{textblock}{5}(9.6,1.5)
\begin{alertblock}{}\fontsize{10}{12}\sf
Principal components based on all features from the feasts package
\end{alertblock}
\end{textblock}

## Feature extraction and statistics
\fontsize{9}{9}\sf

\begin{textblock}{3.3}(.4,3)
\begin{alertblock}{}\fontsize{10}{12}\sf
Principal components based on all features from the feasts package
\end{alertblock}
\end{textblock}

```r
pcs |> ggplot(aes(x=.fittedPC1, y=.fittedPC2)) +
  geom_point() + theme(aspect.ratio=1)
```

\placefig{4}{2.6}{height=6.4cm, width=12cm}{pca1}
\vspace*{10cm}

## Feature extraction and statistics
\fontsize{9}{9}\sf

\begin{textblock}{3.3}(.4,3)
\begin{alertblock}{}\fontsize{10}{12}\sf
Principal components based on all features from the feasts package
\end{alertblock}
\end{textblock}

```r
pcs |> ggplot(aes(x=.fittedPC1, y=.fittedPC2, col=State)) +
  geom_point() + theme(aspect.ratio=1)
```

\placefig{4}{2.6}{height=6.4cm, width=12cm}{pca2}
\vspace*{10cm}

## Feature extraction and statistics
\fontsize{9}{9}\sf

\begin{textblock}{3.3}(.4,3)
\begin{alertblock}{}\fontsize{10}{12}\sf
Principal components based on all features from the feasts package
\end{alertblock}
\end{textblock}

```r
pcs |> ggplot(aes(x=.fittedPC1, y=.fittedPC2, col=Purpose)) +
  geom_point() + theme(aspect.ratio=1)
```

\only<1>{\placefig{4}{2.6}{height=6.4cm, width=12cm}{pca3}}
\only<2>{\placefig{4}{2.6}{height=6.4cm, width=12cm}{pca4}}
\only<3>{\placefig{4}{2.6}{height=6.4cm, width=12cm}{pca5}}
\vspace*{10cm}


# fable: Forecasting

## Model fitting
\fontsize{9}{9}\sf

```{r holiday-model}
training <- tourism |>
  filter(year(Quarter) <= 2015) 
fit <- training |> 
  model(
    snaive = SNAIVE(Trips),
    naive = NAIVE(Trips),
    ets = ETS(Trips),
    arima = ARIMA(Trips)
  )
```

\fontsize{8}{8}\sf

```{r show-models, echo=FALSE, dependson='holiday-model'}
options(width=90)
fit
options(width=80)
```

## Model fitting
\fontsize{9}{10}\sf

```{r holiday-model3, dependson='holiday-model'}
fit |>
  filter(Purpose == "Holiday", Region == "Snowy Mountains") |> 
  select(arima) |>
  report()
```

## Model fitting
\fontsize{9}{10}\sf

```{r holiday-model6, dependson='holiday-model'}
augment(fit)
```

## Ljung-Box test
\fontsize{9}{10}\sf

```{r lbtest, echo=TRUE, fig.height=4, warning = FALSE}
augment(fit) |>
  filter(Purpose == "Holiday", Region == "Snowy Mountains", .model == "arima") |>
  features(.resid, ljung_box, dof = 3, lag = 8)
```

## `gg_tsresiduals()` function
\fontsize{9}{9}\sf

```{r ggtsresiduals, echo=TRUE, warning = FALSE}
fit |>
  filter(Purpose == "Holiday", Region == "Snowy Mountains") |> 
  select(arima) |>
  gg_tsresiduals()
```

## Producing forecasts

\fontsize{9}{9}\sf

```{r holiday-fc, echo = TRUE, dependson='holiday-model'}
fc <- fit |>
  forecast(h = "2 years")
```

```{r holiday-fbl, echo = FALSE, dependson='holiday-fc'}
print(fc, n = 10)
```

## Visualising forecasts
\fontsize{9}{9}\sf

```{r holiday-fc-plot, warning=FALSE, message=FALSE, fig.height=2.9, dependson='holiday-fc'}
fc |>
  filter(Purpose == "Holiday", Region == "Snowy Mountains") |> 
  autoplot(training, level = NULL) +
  labs(title = "Snowy Mountain Holidays", y = "Thousands of visitors") +
  guides(color = guide_legend(title = "Forecast"))
```


## Visualising forecasts
\fontsize{9}{9}\sf

```{r holiday-fc-plot2, warning=FALSE, message=FALSE, fig.height=2.9, dependson='holiday-fc'}
fc |>
  filter(Purpose == "Holiday", Region == "Snowy Mountains", .model == "arima") |>
  autoplot(training) +
  labs(title = "Holidays in Snowy Mountains", y = "Thousands of visitors") +
  guides(color = guide_legend(title = "Forecast"))
```

## Prediction intervals
\fontsize{9}{10}\sf

```{r holiday-fc-interval, dependson='holiday-fc'}
fc |> hilo(level = 95)
```

## Measures of forecast accuracy

```{r holidays-fc-1, echo=FALSE, fig.height=4}
fc |>
  filter(Purpose == "Holiday", Region == "Snowy Mountains") |> 
  autoplot(tourism, level = NULL) +
  labs(title = "Snowy Mountain Holidays", y = "Thousands of visitors") +
  guides(color = guide_legend(title = "Forecast"))
```

## Measures of forecast accuracy
\fontsize{10}{10}\sf

```{r holidays-train-accuracy}
fc |> 
  accuracy(tourism)
```

## Measures of forecast accuracy
\fontsize{10}{10}\sf

```{r holidays-train-accuracy2}
fc |> 
  accuracy(tourism) |> 
  summarise(RMSSE = sqrt(mean(RMSSE^2)), .by=".model")
```

# fable: Forecast reconciliation

## Hierarchical time series
\fontsize{13}{14}\sf

A \alert{\textbf{hierarchical time series}} is a collection of several time series that are linked together in a hierarchical structure.

\begin{minipage}{9.6cm}
\begin{block}{}
\begin{tikzpicture}
\tikzstyle{every node}=[ellipse,draw,inner sep=0.2pt,fill=red!15]
\tikzstyle[level distance=.1cm]
\tikzstyle[sibling distance=7cm]
\tikzstyle{level 1}=[sibling distance=33mm,set style={{every node}+=[fill=blue!15]}]
\tikzstyle{level 2}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=yellow]}]
\node{Total}[edge from parent fork down]
 child {node {A}
   child {node {AA}}
   child {node {AB}}
   child {node {AC}}
 }
 child {node {B}
   child {node {BA}}
   child {node {BB}}
   child {node {BC}}
 }
 child {node {C}
   child {node {CA}}
   child {node {CB}}
   child {node {CC}}
 };
\end{tikzpicture}
\end{block}
\end{minipage}

\pause\alert{Examples}\vspace*{-0.2cm}

 * Tourism demand by states, zones, regions

## Grouped time series
\fontsize{13}{14}\sf

A \alert{\textbf{grouped time series}} is a collection of time series that can be grouped together in a number of non-hierarchical ways.

\begin{minipage}{9.2cm}
\begin{block}{}
\begin{tikzpicture}[level distance=1.5cm]
\tikzstyle{every node}=[ellipse,draw,inner sep=0.2pt,outer sep=0pt, fill=red!15]
\tikzstyle{level 1}=[sibling distance=23mm,set style={{every node}+=[fill=blue!15]},level distance=1cm]
\tikzstyle{level 2}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=yellow]}, level distance=0.9cm]
\node{Total}[edge from parent fork down]
 child {node {A}
   child {node {AX}}
   child {node {AY}}
 }
 child {node {B}
   child {node {BX}}
   child {node {BY}}
 };
\end{tikzpicture}\hspace*{1cm}
\begin{tikzpicture}[level distance=1.5cm]
\tikzstyle{every node}=[ellipse,draw,inner sep=0.2pt,outer sep=0pt, fill=red!15]
\tikzstyle{level 1}=[sibling distance=23mm,set style={{every node}+=[fill=blue!15]},level distance=1cm]
\tikzstyle{level 2}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=yellow]}, level distance=0.9cm]
\node{Total}[edge from parent fork down]
 child {node {X}
   child {node {AX}}
   child {node {BX}}
 }
 child {node {Y}
   child {node {AY}}
   child {node {BY}}
 };
\end{tikzpicture}
\end{block}
\end{minipage}

\pause\alert{Examples}

 * Tourism by state and purpose of travel
 * Retail sales by product groups/sub groups, and by countries/regions

## Creating aggregates
\fontsize{8}{8}\sf

```{r tourism_aggregate}
tourism |>
  aggregate_key(Purpose * (State / Region), Trips = sum(Trips)) |>
  filter(Quarter == yearquarter("1998 Q1")) |>
  print(n = 15)
```

## Creating aggregates
\fontsize{13}{15}\sf

 * Similar to `summarise()` but using the key structure
 * A grouped structure is specified using `grp1 * grp2`
 * A nested structure is specified via `parent / child`.
 * Groups and nesting can be mixed:

    ```r
    (country/region/city) * (brand/product)
    ```

 * All possible aggregates are produced.
 * These are useful when forecasting at different levels of aggregation.


## The problem
\fontsize{13}{14}\sf

\begin{alertblock}{}
\begin{enumerate}\tightlist
 \item How to forecast time series at all nodes such that the forecasts add up in the same way as the original data?
 \item Can we exploit relationships between the series to improve the forecasts?
\end{enumerate}
\end{alertblock}\pause

### The solution

1. Forecast all series at all levels of aggregation using an automatic forecasting algorithm.\newline (e.g., `ETS`, `ARIMA`, ...)
2. Reconcile the resulting forecasts so they add up correctly using least squares optimization (i.e., find closest reconciled forecasts to the original forecasts).
3. This is available using `reconcile()`.

## Forecast reconciliation
\fontsize{9}{10}\sf

```{r tourismets_reconciled, message=FALSE}
tourism |>
  aggregate_key(Purpose * (State / Region), Trips = sum(Trips)) |>
  model(ets = ETS(Trips)) |>
  reconcile(ets_adjusted = min_trace(ets)) |>
  forecast(h = 2)
```

## Hierarchical and grouped time series

Every collection of time series with aggregation constraints can be written as
\begin{block}{}
\centerline{$\by_{t}=\bS\bm{b}_{t}$}
\end{block}
where

 * $\by_t$ is a vector of all series at time $t$
 * $\bm{b}_t$ is a vector of the most disaggregated series at time $t$
 * $\bS$ is a ``summing matrix'' containing the aggregation constraints.

## Hierarchical time series

\begin{minipage}{4cm}\vspace*{0.2cm}
\begin{block}{}\centering
\begin{tikzpicture}
\tikzstyle{every node}=[ellipse,draw,fill=red!15,inner sep=2pt]
\tikzstyle[level distance=.3cm]
\tikzstyle[sibling distance=12cm]
\tikzstyle{level 1}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=blue!15]}]
\node{Total}[edge from parent fork down]
 child {node {A}
 }
 child {node {B}
 }
 child {node {C}
 };
\end{tikzpicture}
\end{block}
\end{minipage}

\only<2->{\begin{textblock}{6.3}(6,1)\small
\begin{itemize}\itemsep=0cm\parskip=0cm
\item[$ y_{t}: $] observed aggregate of all series at time
$t$.
\item[$ y_{X,t}: $] observation on series $X$ at time $t$.
\item[$ \bm{b}_{t}: $] vector of all series at bottom level
in time $t$.
\end{itemize}
\end{textblock}}\vspace*{0.6cm}
\only<3->{
$\bY_{t}= \begin{pmatrix}
  y_{t}\\
  y_{A,t}\\
  y_{B,t}\\
  y_{C,t}
  \end{pmatrix} = \only<3>{\hspace*{0.01cm}\begin{pmatrix}
                1 & 1 & 1 \\
                1 & 0 & 0 \\
                0 & 1 & 0\\
                0 & 0 & 1
                \end{pmatrix}}\only<4->{{\color{blue}\underbrace{\begin{pmatrix}
                1 & 1 & 1 \\
                1 & 0 & 0 \\
                0 & 1 & 0\\
                0 & 0 & 1
                \end{pmatrix}}_{\bS}}}\only<3>{\hspace*{0.08cm}}\only<3>{\hspace*{-0.1cm}\begin{pmatrix}y_{A,t}\\y_{B,t}\\y_{C,t}\end{pmatrix}}\rule{0cm}{1.6cm}
                \only<4->{\hspace*{0.08cm}{\color{red}\underbrace{\begin{pmatrix}y_{A,t}\\y_{B,t}\\y_{C,t}\end{pmatrix}}_{\bm{b}_{t}}}}$}

\vspace*{-0.6cm}

\only<4>{\hspace*{8cm}\colorbox[RGB]{210,210,210}{$\bY_{t}=\color{blue}\bS\color{red}\bm{b}_{t}$}}

\vspace*{10cm}

## Forecasting notation

Let $\hat{\by}_n(h)$ be vector of initial $h$-step forecasts, made at time $n$, stacked in same order as $\by_t$. \pause\newline  (In general, they will not ``add up''.)\pause

\begin{block}{}
Reconciled forecasts must be of the form:
$$\tilde{\by}_{n}(h)=\bS\bm{G}\hat{\by}_{n}(h)$$
for some matrix $\bm{G}$.
\end{block}\pause

 * $\bm{G}$ extracts and combines base forecasts $\hat{\by}_{n}(h)$ to get bottom-level forecasts.
 * $\bS$ adds them up

## Optimal combination forecasts
\fontsize{13}{14}\sf

\begin{alertblock}{Main result}
The best (minimum sum of variances) unbiased forecasts are obtained when
$\bm{G} = (\bS'\bW^{-1}_{h}\bS)^{-1}\bS'\bW^{-1}_{h}$,
where $\bW_h$ is the $h$-step base forecast error covariance matrix.
\end{alertblock}

\pause

\begin{block}{}
\centerline{$\displaystyle\textcolor{red}{\tilde{\by}_{n}(h)}
=\bS(\bS'\bW^{-1}_{h}\bS)^{-1}\bS'\bW^{-1}_{h}\textcolor{blue}{\hat{\by}_{n}(h)}$}
\end{block}\fontsize{14}{15}\sf

\alert{\textbf{Problem:}} $\bW_h$ hard to estimate, especially for $h>1$.

\alert{Solutions:}\vspace*{-0.4cm}

 * Ignore $\bW_h$ (OLS) [`min_trace(method='ols')`]
 * Assume $\bW_h = k_h\bW_1$ is diagonal (WLS) [`min_trace(method='wls')`]
 * Assume $\bW_h = k_h\bW_1$ and estimate it (GLS) [`min_trace(method='shrink')` (the default)]

## Example: Australian tourism
\fontsize{11}{11}\sf

```{r fctourism}
tourism_agg <- tourism |>
  aggregate_key(Purpose * (State / Region),
    Trips = sum(Trips)
  )
fc <- tourism_agg |>
  filter(year(Quarter) <= 2015) |>
  model(
    ets = ETS(Trips),
    arima = ARIMA(Trips)
  ) |>
  mutate(
    comb = (ets + arima) / 2
  ) |>
  reconcile(
    ets_adj = min_trace(ets),
    arima_adj = min_trace(arima),
    comb_adj = min_trace(comb)
  ) |>
  forecast(h = "2 years")
```

## Forecast evaluation
\fontsize{10}{11}\sf

```{r fcaccuracy, dependson='fctourismcomb'}
accuracy(fc, tourism_agg)
```

## Forecast evaluation
\fontsize{11}{11}\sf

```{r fcaccuracy2, dependson='fctourismcomb'}
accuracy(fc, tourism_agg) |>
  summarise(RMSSE = sqrt(mean(RMSSE^2)), by=".model") |>
  arrange(RMSSE)
```

## More information
\fontsize{18}{20}\sf

 * Slides and papers: **robjhyndman.com**
 * Packages: **tidyverts.org**
 * Forecasting textbook using tsibble, feasts and fable packages: **OTexts.com/fpp3**

\begin{textblock}{8}(7.6,4.8)
\begin{alertblock}{Find me at \dots}
\href{https://twitter.com/robjhyndman}{\faicon{twitter} @robjhyndman}

\href{https://github.com/robjhyndman}{\faicon{github} @robjhyndman}

\href{https://robjhyndman.com}{\faicon{home} robjhyndman.com}

\href{mailto:rob.hyndman@monash.edu}{\faicon{envelope} rob.hyndman@monash.edu}
\end{alertblock}
\end{textblock}
\vspace*{10cm}
