---
title: "Feasts & fables: modern tools for time series analysis"
author: "Rob J Hyndman"
date: "10 December 2021"
toc: true
classoption: aspectratio=169
output:
  binb::monash:
    colortheme: monashwhite
    highlight: pygments
    fig_width: 8
    fig_height: 3.3
    keep_tex: no
    includes:
      in_header: whyRheader.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = TRUE, warning = FALSE, cache = TRUE,
  dev.args = list(pointsize = 9)
)
```

```{r packages, echo=FALSE, include=FALSE}
# Load packages
library(fpp3)
library(patchwork)
library(glue)
library(stringr)
library(broom)

# Colors to be viridis for continuous scales and Okabe for discrete scales
options(
  digits = 3,
  width = 90,
  ggplot2.continuous.color = "viridis",
  ggplot2.continuous.fill = "viridis",
  ggplot2.discrete.colour = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442"),
  ggplot2.discrete.fill = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442")
)

# Neater data sets
global_economy <- global_economy %>%
  select(Year, Country, GDP, Imports, Exports, Population)
tourism <- tsibble::tourism %>%
  mutate(
    State = recode(State,
      "Australian Capital Territory" = "ACT",
      "New South Wales" = "NSW",
      "Northern Territory" = "NT",
      "Queensland" = "QLD",
      "South Australia" = "SA",
      "Tasmania" = "TAS",
      "Victoria" = "VIC",
      "Western Australia" = "WA"
    ),
    Region = stringr::str_remove(Region, "Australia's ")
  )
```

## Tidyverts packages

\begin{textblock}{3.8}(11,0)\begin{alertblock}{}\Large\textbf{tidyverts.org}\end{alertblock}\end{textblock}

\placefig{3}{1.6}{width=3.5cm}{tsibble.png}
\placefig{6.5}{1.6}{width=3.5cm}{tsibbledata.png}
\placefig{4.75}{4.6}{width=3.5cm}{feasts.png}
\placefig{8.25}{4.6}{width=3.5cm}{fable.png}

# What does modern time series data look like?


## Australian electricity demand

\fontsize{10}{11.2}\sf

```{r, echo=FALSE}
bind_rows(
    as_tibble(vic_elec) %>% mutate(State = "VIC"),
    as_tibble(vic_elec) %>% mutate(State = "NSW"),
    as_tibble(vic_elec) %>% mutate(State = "SA"),
    as_tibble(vic_elec) %>% mutate(State = "WA"),
    as_tibble(vic_elec) %>% mutate(State = "TAS"),
    as_tibble(vic_elec) %>% mutate(State = "NT"),
    as_tibble(vic_elec) %>% mutate(State = "ACT"),
    as_tibble(vic_elec) %>% mutate(State = "QLD")
  ) %>%
  mutate(State = factor(State, levels=c("VIC","NSW","SA","WA","TAS","NT","QLD","ACT"))) %>%
  as_tsibble(index=Time, key = State) %>%
  select(Time, State, Date, Holiday, Temperature, Demand)
```

\only<2->{\begin{textblock}{1.1}(2.1,2.85)
\begin{alertblock}{}\fontsize{10}{10}\sf Index\phantom{dg}\end{alertblock}
\end{textblock}}
\only<3->{\begin{textblock}{0.8}(5.65,2.85)
\begin{alertblock}{}\fontsize{10}{10}\sf Key\phantom{dg}\end{alertblock}
\end{textblock}}
\only<4>{\begin{textblock}{6.5}(6.85,2.85)
\begin{alertblock}{}\fontsize{10}{10}\sf Measures\phantom{dg}\end{alertblock}
\end{textblock}}
## Australian tourism regions

```{r ausmap, echo=FALSE, message=FALSE, width=20, height=18, out.height="90%"}
library(sf)
# Use Okabe-Ito color-blind friendly color palette
state_colors <- c(
  `New South Wales` = "#56b4e9",
  `Victoria` = "#0072b2",
  `Queensland` = "#009e73",
  `South Australia` = "#f0e442",
  `Northern Territory` = "#d55e00",
  `Western Australia` = "#e69f00",
  `Tasmania` = "#cc79a7",
  `Australian Capital Territory` = "#cccccc"
)
read_sf("tourism/Tourism_Regions_2020.shp") %>%
  rename(State = "STE_NAME16") %>%
  ggplot() +
  geom_sf(aes(fill = State), alpha = 0.8) +
  theme_void() +
  scale_fill_manual(values = state_colors)
```

\only<2>{\begin{textblock}{6.5}(9.1,1.4)
\begin{block}{}\fontsize{12}{12}\sf
  \begin{itemize}\itemsep=0cm\parskip=0cm
    \item Quarterly data on visitor nights: 1998 -- 2017
    \item From \textit{National Visitor Survey}, interviews of 120,000 Australians aged 15+.
    \item Geographical hierarchy split by
    \begin{itemize}
    \item 8 states and territories
    \item 76 regions
    \end{itemize}
    \item Purpose:
    \begin{itemize}
    \item Holidays
    \item Business
    \item Visiting friends \& relatives
    \item Other
    \end{itemize}
  \end{itemize}
\end{block}
\end{textblock}}

## Quarterly tourism data

\fontsize{10}{11.3}\sf

```{r, echo = FALSE}
tourism
```

\only<2->{\begin{textblock}{1.1}(2.1,2.85)
\begin{alertblock}{}\fontsize{10}{10}\sf Index\phantom{dg}\end{alertblock}
\end{textblock}}
\only<3->{\begin{textblock}{3.9}(3.65,2.85)
\begin{alertblock}{}\fontsize{10}{10}\sf Keys\phantom{dg}\end{alertblock}
\end{textblock}}
\only<4>{\begin{textblock}{1.5}(7.95,2.85)
\begin{alertblock}{}\fontsize{10}{10}\sf Measure\phantom{dg}\end{alertblock}
\end{textblock}}

\begin{textblock}{3}(9,5)
\begin{block}{}\fontsize{10}{10}\sf Domestic visitor nights in thousands by state/region and purpose.\phantom{dg}\end{block}
\end{textblock}

## Characteristics of modern time series

  * Often observed at sub-daily frequency over a long time.
  * Multiple keys which may be nested.
  * Multiple seasonal patterns.
  * Multiple measures for each combination of index and keys.
  
\pause\vspace*{0.3cm}

### `tsibble` objects
\fontsize{12}{13}\sf

  * A `tsibble` allows storage and manipulation of multiple time series in R.
  * It contains:

    + An index: time information about the observation
    + Key variable(s): optional unique identifiers for each series
    + Measured variable(s): numbers of interest and any other variable

# Feature-based time series analysis

## STL decomposition

\begin{textblock}{5}(9,0.2)
\begin{alertblock}{STL decomposition}
\centerline{$y_t = T_t+S_t+R_t$}
\end{alertblock}
\end{textblock}
\begin{textblock}{1}(8.6,2.6)
$y_t$\\[.7cm]
$T_t$\\[.7cm]
$S_t$\\[.7cm]
$R_t$
\end{textblock}

```{r, echo=FALSE, fig.height=6, fig.width=7}
tourism %>%
  filter(Purpose == "Holiday", Region == "Snowy Mountains") %>%
  model(STL(Trips ~ trend(window = 13))) %>%
  components() %>%
  autoplot() +
  labs(title = "STL decomposition: Holidays in Snowy Mountains",
       y = "Overnight trips (thousands)")
```

\begin{textblock}{6}(9.5,2.6)
\only<2->{\begin{block}{Trend strength}
$$\max\left(0, 1-\frac{\text{Var}(R_t)}{\text{Var}(T_t+R_t)}\right)$$
\end{block}}
\only<3>{\begin{block}{Seasonal strength}
$$\max\left(0, 1-\frac{\text{Var}(R_t)}{\text{Var}(S_t+R_t)}\right)$$
\end{block}}
\end{textblock}

## STL-based features
\fontsize{9}{9}\sf

```{r features}
tourism %>%
  features(Trips, feat_stl)
```

## STL-based features
\fontsize{9}{9}\sf

```{r features-plot, fig.height=4, fig.width=6, out.height="67%"}
tourism %>%
  features(Trips, feat_stl) %>%
  ggplot(aes(x = trend_strength, y = seasonal_strength_year, col = Purpose)) +
  geom_point() + facet_wrap(vars(State))
```

\only<2->{\begin{textblock}{5.3}(9.8,6.6)
\begin{alertblock}{}\fontsize{12}{12}\sf
\begin{itemize}\tightlist
\item Holidays more seasonal than other travel.
\item WA has strongest trends.
\end{itemize}
\end{alertblock}\end{textblock}}

## STL-based features

Find the most seasonal time series:

\fontsize{9}{9}\sf

```{r extreme, fig.height=2}
tourism %>%
  features(Trips, feat_stl) %>%
  filter(seasonal_strength_year == max(seasonal_strength_year)) %>%
  left_join(tourism, by = c("State", "Region", "Purpose")) %>%
  ggplot(aes(x = Quarter, y = Trips)) +
  geom_line() +
  facet_grid(vars(State, Region, Purpose))
```

## STL-based features

Find the most trended time series:

\fontsize{9}{9}\sf

```{r extreme3, fig.height=2}
tourism %>%
  features(Trips, feat_stl) %>%
  filter(trend_strength == max(trend_strength)) %>%
  left_join(tourism, by = c("State", "Region", "Purpose")) %>%
  ggplot(aes(x = Quarter, y = Trips)) +
  geom_line() +
  facet_grid(vars(State, Region, Purpose))
```

```{r pca, echo=FALSE}
# Save pdf figures
savepdf <- function(file, width = 16, height = 10) {
  fname <<- paste("figs/", file, ".pdf", sep = "")
  pdf(fname, width = width / 2.54, height = height / 2.54, pointsize = 10)
  par(mgp = c(2.2, 0.45, 0), tcl = -0.4, mar = c(3.3, 3.6, 1.1, 1.1))
}
endpdf <- function() {
  crop::dev.off.crop(fname)
}
# Compute features
tourism_features <- tourism %>%
  features(Trips, feature_set(pkgs = "feasts"))
# Compute PCs
pcs <- tourism_features %>%
  select(-State, -Region, -Purpose) %>%
  prcomp(scale = TRUE) %>%
  broom::augment(tourism_features)
# Save some PC plots
savepdf("pca1", 18, 10)
pcs %>% ggplot(aes(x = .fittedPC1, y = .fittedPC2)) +
  geom_point() +
  theme(aspect.ratio = 1)
endpdf()
savepdf("pca2", 18, 10)
pcs %>% ggplot(aes(x = .fittedPC1, y = .fittedPC2, col = State)) +
  geom_point() +
  theme(aspect.ratio = 1)
endpdf()
savepdf("pca3", 18, 10)
pcs %>% ggplot(aes(x = .fittedPC1, y = .fittedPC2, col = Purpose)) +
  geom_point() +
  theme(aspect.ratio = 1)
endpdf()
# Find outliers
outliers <- pcs %>%
  filter(.fittedPC1 > 10, .fittedPC2 > 0, .fittedPC2 < 5)
savepdf("pca4", 18, 10)
pcs %>% ggplot(aes(x = .fittedPC1, y = .fittedPC2, col = Purpose)) +
  geom_point() +
  theme(aspect.ratio = 1) +
  geom_point(data = outliers, aes(x = .fittedPC1, y = .fittedPC2), col = "black", shape = 1, size = 3)
endpdf()
```

## Time series features
\fontsize{9}{9}\sf

```{r tourismfeatures, eval=FALSE}
tourism_features <- tourism %>%
  features(Trips, feature_set(pkgs = "feasts"))
```

\vspace*{-0.5cm}\fontsize{8}{8}\sf

```{r echo=FALSE}
tourism_features
```

\begin{textblock}{5}(9.6,1.5)
\begin{alertblock}{}\fontsize{10}{12}\sf
All features from the feasts package
\end{alertblock}
\end{textblock}

## Reduced feature space
\fontsize{9}{9}\sf

```{r pcatable}
pcs <- tourism_features %>%
  select(-State, -Region, -Purpose) %>%
  prcomp(scale = TRUE) %>%
  augment(tourism_features)
```

\vspace*{-0.5cm}\fontsize{8}{8}\sf

```{r echo=FALSE}
pcs
```

\begin{textblock}{5}(9.6,1.5)
\begin{alertblock}{}\fontsize{10}{12}\sf
Principal components based on all features from the feasts package
\end{alertblock}
\end{textblock}

## Reduced feature space
\fontsize{9}{9}\sf

\begin{textblock}{3.3}(.4,3)
\begin{alertblock}{}\fontsize{10}{12}\sf
Principal components based on all features from the feasts package
\end{alertblock}
\end{textblock}

```r
pcs %>% ggplot(aes(x=.fittedPC1, y=.fittedPC2)) +
  geom_point() + theme(aspect.ratio=1)
```

\placefig{4}{2.6}{height=6.4cm, width=12cm}{pca1}
\vspace*{10cm}

## Reduced feature space
\fontsize{9}{9}\sf

\begin{textblock}{3.3}(.4,3)
\begin{alertblock}{}\fontsize{10}{12}\sf
Principal components based on all features from the feasts package
\end{alertblock}
\end{textblock}

```r
pcs %>% ggplot(aes(x=.fittedPC1, y=.fittedPC2, col=State)) +
  geom_point() + theme(aspect.ratio=1)
```

\placefig{4}{2.6}{height=6.4cm, width=12cm}{pca2}
\vspace*{10cm}

## Reduced feature space
\fontsize{9}{9}\sf

\begin{textblock}{3.3}(.4,3)
\begin{alertblock}{}\fontsize{10}{12}\sf
Principal components based on all features from the feasts package
\end{alertblock}
\end{textblock}

```r
pcs %>% ggplot(aes(x=.fittedPC1, y=.fittedPC2, col=Purpose)) +
  geom_point() + theme(aspect.ratio=1)
```

\placefig{4}{2.6}{height=6.4cm, width=12cm}{pca3}
\vspace*{10cm}

## Anomaly detection using time series features
\fontsize{9}{9}\sf

\begin{textblock}{3.3}(.4,3)
\begin{alertblock}{}\fontsize{10}{12}\sf
Principal components based on all features from the feasts package
\end{alertblock}
\end{textblock}

```r
pcs %>% ggplot(aes(x=.fittedPC1, y=.fittedPC2, col=Purpose)) +
  geom_point() + theme(aspect.ratio=1)
```

\placefig{4}{2.6}{height=6.4cm, width=12cm}{pca4}
\vspace*{10cm}

## Anomaly detection using time series features
\fontsize{8}{8}\sf

```{r outliers2, fig.height=6, fig.width=10, out.width="100%", dependson='pca', echo=FALSE}
out <- outliers %>%
  left_join(tourism, by = c("State", "Region", "Purpose")) %>%
  mutate(
    Series = glue("{State}", "{Region}", "{Purpose}", .sep = "\n\n"),
    Series = as.character(Series)
  ) %>%
  select(Quarter, Series, Trips, .fittedPC2)
ordering <- out %>% 
  select(Series,.fittedPC2) %>%
  arrange(desc(.fittedPC2)) %>% 
  select(Series) %>%
  distinct(Series) %>%
  pull(Series)
out %>%
  mutate(Series = factor(Series, levels=ordering)) %>%
  select(Series, Quarter, Trips) %>%
  ggplot(aes(x = Quarter, y = Trips)) + geom_line() +
  facet_grid(Series ~ ., scales = "free_y") +
  ggtitle("Outlying time series in PC space")
```

# Probabilistic forecasting for large time series

```{r austa, include=FALSE}
austa <- tourism %>% summarise(Trips = sum(Trips)/1e3)
# Fit ETS model
fit <- austa %>%
  model(ETS(Trips ~ trend("A")))
# Product forecasts
fc <- forecast(fit, h = 16)
# Simulate 100 future sample paths
set.seed(1967)
sim <- fit %>%
  generate(h = 16, times = 100) %>%
  mutate(
    replicate = factor(.rep, levels = 1:100, labels = paste("Future", 1:100)),
    .rep = as.numeric(.rep)
  ) %>%
  as_tibble()
# Nice colors
cols <- scale_color_manual(
  values = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442", sample(rainbow(93))),
  breaks = paste("Future", 1:100),
  name = " "
)
# Now build up some plots with alignment
p1 <- austa %>%
  bind_rows(new_data(austa, 16)) %>%
  as_tibble() %>%
  ggplot(aes(x = Quarter, y = Trips)) +
  geom_line() +
  labs(
    x = "Quarter",
    y = "Overnight trips (millions)",
    title = "Total Australian domestic overnight trips"
  ) +
  scale_x_yearquarter(
    breaks = yearquarter(seq(as.Date("2000-01-01"), by = "5 years", l = 5)),
    labels = paste(seq(2000, 2020, by = 5),"Q1"),
    minor_breaks = yearquarter(seq(as.Date("2000-01-01"), by = "1 year", l = 25)),
  ) +
  ylim(min(austa$Trips, sim$.sim, na.rm = TRUE), max(austa$Trips, sim$.sim, na.rm = TRUE))
p2 <- p1 + cols +
  geom_line(aes(y = .sim, group = replicate, color = replicate),
    data = sim %>% filter(.rep <= 1)
  )
p3 <- p1 + cols +
  geom_line(aes(y = .sim, group = replicate, color = replicate),
    data = sim %>% filter(.rep <= 2)
  )
p4 <- p1 + cols +
  geom_line(aes(y = .sim, group = replicate, color = replicate),
    data = sim %>% filter(.rep <= 3)
  )
p5 <- p1 + cols +
  geom_line(aes(y = .sim, group = replicate, color = replicate),
    data = sim %>% filter(.rep <= 10)
  )
p6 <- p1 + cols +
  geom_line(aes(y = .sim, group = replicate, color = replicate), alpha = 0.5, data = sim) +
  guides(color = "none")
p7 <- p1 +
  geom_line(aes(y = .sim, group = replicate, color = replicate), alpha = 0.5, col = "gray", data = sim) +
  guides(color = "none")
p8 <- p7 + autolayer(fc, level = c(50, 90))
p9 <- p8 + coord_cartesian(xlim = as.Date(c("2013-01-01", "2021-08-01"))) +
  scale_x_yearquarter(
    breaks = yearquarter(seq(as.Date("2000-01-01"), by = "2 years", l = 11)),
    labels = paste(seq(2000, 2020, by = 2),"Q1"),
    minor_breaks = yearquarter(seq(as.Date("2000-01-01"), by = "1 year", l = 25)),
  )
aligned_plots <- align_patches(p1, p2, p3, p4, p5, p6, p7, p8, p9)

for (i in seq(9)) {
  pdf(paste0("figs/austa", i, ".pdf"), height = 3.3, width = 8)
  print(aligned_plots[[i]])
  dev.off()
}
```

## Random futures

\forecast\pause
\includegraphics[height=8cm]{austa1}

## Random futures

\forecast
\includegraphics[height=8cm]{austa2}
\simfutures

## Random futures

\forecast
\includegraphics[height=8cm]{austa3}
\simfutures

## Random futures

\forecast
\includegraphics[height=8cm]{austa4}
\simfutures

## Random futures

\forecast
\includegraphics[height=8cm]{austa5}
\simfutures

## Random futures

\forecast
\includegraphics[height=8cm]{austa6}
\simfutures

## Random futures

\forecast
\includegraphics[height=8cm]{austa7}
\simfutures

## Random futures

\forecast
\includegraphics[height=8cm]{austa8}
\simfutures

## Random futures

\forecast
\includegraphics[height=8cm]{austa9}
\simfutures

## Model fitting
\fontsize{9}{9}\sf

```{r holiday-model}
tourism_fit <- tourism %>%
  filter(year(Quarter) <= 2015) %>%
  model(ets = ETS(Trips), arima = ARIMA(Trips)) %>%
  mutate(ensemble = (ets + arima)/2)
```

```{r holiday-model2, echo=FALSE, dependson='holiday-model'}
tourism_fit
```

## Producing forecasts

\fontsize{9}{9}\sf

```{r holiday-fc, echo = TRUE, dependson='holiday-model'}
tourism_fc <- tourism_fit %>%
  forecast(h = "2 years")
```

```{r holiday-fbl, echo = FALSE, dependson='holiday-fc'}
print(tourism_fc, n = 10)
```

## Visualising forecasts
\fontsize{9}{9}\sf

```{r fcviz, include=FALSE, dependson='holiday-fc'}
p1 <- tourism_fc %>%
  filter(Region == "Adelaide", Purpose=="Business") %>%
  autoplot(tourism %>% filter(year(Quarter) <= 2015), level = NULL) +
  labs(title = "Adelaide Business Trips", y = "Thousands of visitors") +
  guides(color = guide_legend(title = "Forecast"))
p2 <- tourism_fc %>%
  filter(Region == "Adelaide", Purpose=="Business", .model == "arima") %>%
  autoplot(tourism %>% filter(year(Quarter) <= 2015)) +
  labs(title = "Adelaide Business Trips", y = "Thousands of visitors") +
  guides(color = guide_legend(title = "Forecast"))
p3 <- p2 +
  geom_line(aes(x=Quarter, y=Trips), data = tourism %>% filter(Region=="Adelaide", Purpose=="Business"))
aligned_plots <- align_patches(p1,p2,p3)
for (i in seq(3)) {
  pdf(paste0("figs/austafc", i, ".pdf"), height = 2.9, width = 8)
  print(aligned_plots[[i]])
  dev.off()
}
```

```r
tourism_fc %>%
  filter(Region == "Adelaide", Purpose=="Business") %>%
  autoplot(tourism, level = NULL) +
  labs(title = "Adelaide Business Trips", y = "Thousands of visitors") +
  guides(color = guide_legend(title = "Forecast"))
```

\includegraphics[height=8cm]{austafc1}

## Visualising forecasts
\fontsize{9}{9}\sf

```r
tourism_fc %>%
  filter(Region == "Adelaide", Purpose=="Business", .model == "arima") %>%
  autoplot(tourism) +
  labs(title = "Adelaide Business Trips", y = "Thousands of visitors") +
  guides(color = guide_legend(title = "Forecast"))
```

\only<1>{\includegraphics[height=8cm]{austafc2}}
\only<2>{\includegraphics[height=8cm]{austafc3}}

# Evaluating probabilistic forecasts

## Evaluating probabilistic forecasts

```{r traintest, fig.height=1, echo=FALSE, cache=TRUE}
train <- 1:18
test <- 19:24
par(mar = c(0, 0, 0, 0))
plot(0, 0, xlim = c(0, 26), ylim = c(0, 2), xaxt = "n", yaxt = "n", bty = "n", xlab = "", ylab = "", type = "n")
arrows(0, 0.5, 25, 0.5, 0.05)
points(train, train * 0 + 0.5, pch = 19, col = "#0072B2")
points(test, test * 0 + 0.5, pch = 19, col = "#D55E00")
text(26, 0.5, "time")
text(10, 1, "Training data", col = "#0072B2")
text(21, 1, "Test data", col = "#D55E00")
```

```{r tscvplots, echo=FALSE}
tscv_plot <- function(.init, .step, h = 1) {
  expand.grid(
    time = seq(26),
    .id = seq(trunc(20 / .step))
  ) %>%
    group_by(.id) %>%
    mutate(
      observation = case_when(
        time <= ((.id - 1) * .step + .init) ~ "train",
        time %in% c((.id - 1) * .step + .init + h) ~ "test",
        TRUE ~ "unused"
      )
    ) %>%
    ungroup() %>%
    filter(.id <= 26 - .init) %>%
    ggplot(aes(x = time, y = .id)) +
    geom_segment(
      aes(x = 0, xend = 27, y = .id, yend = .id),
      arrow = arrow(length = unit(0.015, "npc")),
      col = "black", size = .25
    ) +
    geom_point(aes(col = observation), size = 2) +
    scale_y_reverse() +
    scale_color_manual(values = c(train = "#0072B2", test = "#D55E00", unused = "gray")) +
    # theme_void() +
    # geom_label(aes(x = 28.5, y = 1, label = "time")) +
    guides(col = FALSE) +
    labs(x = "time", y = "") +
    theme_void() +
    theme(axis.title = element_text())
}
```

```{r qs, include=FALSE}
# Run to generate the graphs
qp <- function(p, y) {
  if (length(p) > 1 & length(y) > 1) {
    stop("Either p or y should be a scalar")
  }
  fp <- qgamma(p, 2)
  2 * (1 - p) * abs(y - fp) * (y < fp) + 2 * p * abs(y - fp) * (y >= fp)
}
gamma <- tibble(
  y = seq(0, 10, l = 101),
  f = dgamma(y, 2)
)
actual <- 3
```

\vspace*{1cm}

```{r pdf1, dependson='qs', echo=FALSE}
# Single pdf graph for median
prob <- 0.5
qactual <- qgamma(prob, 2)
gamma %>%
  ggplot(aes(x = y, y = f)) +
  geom_line() +
  geom_point(aes(x = actual, y = y), data = tibble(actual = actual, y = 0), col="#D55E00") +
  geom_label(aes(x = actual, y = 0.02, label = "Actual"), col="#D55E00") +
  labs(y = "Forecast probability density", x = latex2exp::TeX("Quantile $q_{p}$")) +
  geom_line(aes(x = y, y = f),
    col = "blue", linetype = 2,
    data = tibble(y = rep(qactual, 2), f = c(dgamma(qactual, 2), 0))
  ) +
  geom_label(aes(x = qactual, y = 0.02, label = "Median"), col = "blue") +
  geom_label(aes(x = qactual, y = -0.01, label = "0.5 quantile"), col = "blue") +
  coord_cartesian(ylim = c(-0, 0.375), xlim = c(0, 10), expand = FALSE, clip = "off")
```

## Evaluating probabilistic forecasts

```{r ref.label="traintest", fig.height=1, echo=FALSE, cache=TRUE}
```

\vspace*{1cm}

```{r pdf2, dependson='qs', echo=FALSE}
# Single pdf graph for 0.8 quantile
prob <- 0.9
qactual <- qgamma(prob, 2)
gamma %>%
  ggplot(aes(x = y, y = f)) +
  geom_line() +
  geom_point(aes(x = actual, y = y), data = tibble(actual = actual, y = 0), col="#D55E00") +
  geom_label(aes(x = actual, y = 0.02, label = "Actual"), col="#D55E00") +
  labs(y = "Forecast probability density", x = latex2exp::TeX("Quantile $q_{p}$")) +
  geom_line(aes(x = y, y = f),
    col = "blue", linetype = 2,
    data = tibble(y = rep(qactual, 2), f = c(dgamma(qactual, 2), 0))
  ) +
  geom_label(aes(x = qactual, y = -0.01, label = paste(sprintf("%.2f", prob), "quantile")), col = "blue") +
  coord_cartesian(ylim = c(-0, 0.375), xlim = c(0, 10), expand = FALSE, clip = "off")
```

## Evaluating probabilistic forecasts

```{r pdfqs, dependson='qs', echo=FALSE}
# pdf plus quantile score graph for several quantiles
prob <- c(0.1, 0.3, 0.5, 0.7, 0.9)
qactual <- qgamma(prob, 2)
qpactual <- qp(prob, actual)
for (i in seq_along(prob)) {
  p1 <- gamma %>%
    ggplot(aes(x = y, y = f)) +
    geom_line() +
    geom_point(aes(x = actual, y = y), data = tibble(actual = actual, y = 0), col="#D55E00") +
    labs(y = "Forecast probability density", x = latex2exp::TeX("Quantile $q_{p}$")) +
    geom_line(aes(x = y, y = f),
      col = "blue", linetype = 2,
      data = tibble(y = rep(qactual[i], 2), f = c(dgamma(qactual[i], 2), 0))
    ) +
    geom_label(aes(x = actual, y = 0.03, label = "Actual"), col="#D55E00") +
    geom_label(aes(x = qactual[i], y = -0.02, label = paste(sprintf("%.2f", prob[i]), "quantile")), col = "blue") +
    coord_cartesian(ylim = c(-0, 0.375), xlim = c(0, 10), expand = FALSE, clip = "off")
  p2 <- tibble(
    y = seq(0, 10, l = 101),
    Q = qp(prob[i], y)
  ) %>%
    ggplot(aes(x = y, y = Q)) +
    geom_line() +
    labs(
      x = latex2exp::TeX("Quantile $q_{p}$"),
      y = latex2exp::TeX("Quantile score $S_{p}$")
    ) +
    geom_line(aes(x = y, y = Q),
      col = "red", linetype = 2,
      data = tibble(y = c(actual, actual), Q = c(14.5, qpactual[i]))
    ) +
    geom_line(aes(x = y, y = Q),
      col = "red", linetype = 2,
      data = tibble(y = c(actual, 0), Q = rep(qpactual[i], 2))
    ) +
    geom_line(aes(x = y, y = f),
      col = "blue", linetype = 2,
      data = tibble(y = rep(qactual[i], 2), f = c(0, 13.5))
    ) +
    coord_cartesian(ylim = c(0, 11.6), xlim = c(0, 10), expand = FALSE, clip = "off")
  pdf(paste0("figs/pdfqs", i, ".pdf"), width = 15 / 1.5, height = 8 / 1.5)
  print(p1 / p2)
  crop::dev.off.crop()
  if(i==5) {
    pdf(paste0("figs/pdfqs", i, "a.pdf"), width = 14.9 / 1.5, height = 4 / 1.5)
    print(p1)
    crop::dev.off.crop()
  }
}
```

\only<7>{\placefig{0.2}{1.4}{width=14.7cm,height=20cm}{pdfqs1}}
\only<6>{\placefig{0.2}{1.4}{width=14.7cm,height=20cm}{pdfqs2}}
\only<5>{\placefig{0.2}{1.4}{width=14.7cm,height=20cm}{pdfqs3}}
\only<4>{\placefig{0.2}{1.4}{width=14.7cm,height=20cm}{pdfqs4}}
\only<3>{\placefig{0.2}{1.4}{width=14.7cm,height=20cm}{pdfqs5}}
\only<1-2>{\placefig{0.4}{1.5}{width=14.4cm,height=20cm}{pdfqs5a}}

\only<2->{\begin{textblock}{5.8}(9.8,.4)\fontsize{11}{12}\sf
\begin{alertblock}{}\vspace*{-0.3cm}
\begin{align*}
q_{p} &= \text{quantile forecast with prob. $p$}\\
y &= \text{observation}
\end{align*}
\end{alertblock}\vspace*{-0.3cm}
\begin{alertblock}{Quantile score}\vspace*{-0.1cm}
$$
  S_p(y) = \begin{cases}
  2(1 - p) \big|y - q_{p}\big|, & \text{if $y < q_{p}$}\\
  2p \big|y - q_{p}\big|, & \text{if $y \ge q_{p}$} \end{cases}
$$
\end{alertblock}
\end{textblock}}


## Evaluating probabilistic forecasts

```{r pdfqs2, include=FALSE, dependson='qs'}
qs <- seq(0, 10, l = 200)
prob <- pgamma(qs, 2)
prob <- c(prob[prob <= 0.999], 0.9999)
for (i in seq_along(prob)) {
  qactual <- qgamma(prob[i], 2)
  qpactual <- qp(prob[i], actual)
  p1 <- gamma %>%
    ggplot(aes(x = y, y = f)) +
    geom_line() +
    geom_point(aes(x = actual, y = y), data = tibble(actual = actual, y = 0), col = "#D55E00") +
    geom_label(aes(x = actual, y = 0.02, label = "Actual"), col = "#D55E00") +
    labs(y = "Forecasting probability density", x = latex2exp::TeX("Quantile $q_{p}$")) +
    geom_line(aes(x = y, y = f),
      col = "blue", linetype = 2,
      data = tibble(y = rep(qactual, 2), f = c(dgamma(qactual, 2), -0.1))
    ) +
    coord_cartesian(ylim = c(-0, 0.375), xlim = c(0, 10), expand = FALSE, clip = "off")
  p2 <- tibble(
    y = seq(0, 10, l = 101),
    Q = qp(prob[i], y)
  ) %>%
    ggplot(aes(x = y, y = Q)) +
    geom_line() +
    labs(
      x = latex2exp::TeX("Quantile $q_{p}$"),
      y = latex2exp::TeX("Quantile score $S_{p}$")
    ) +
    geom_line(aes(x = y, y = Q),
      col = "red", linetype = 2,
      data = tibble(y = rep(actual, 2), Q = c(14.5, qpactual))
    ) +
    geom_line(aes(x = y, y = Q),
      col = "red", linetype = 2,
      data = tibble(y = c(actual, -1), Q = rep(qpactual, 2))
    ) +
    geom_line(aes(x = y, y = f),
      col = "blue", linetype = 2,
      data = tibble(y = rep(qactual, 2), f = c(0, 12))
    ) +
    geom_label(aes(x = qactual, y = 11.6, label = paste(sprintf("%.3f", prob[i]), "quantile")), col = "blue") +
    coord_cartesian(ylim = c(0, 11.6), xlim = c(0, 10), expand = FALSE, clip = "off")

  p3data <- tibble(
    p = prob,
    Q = qp(prob, actual)
  ) %>%
    filter(p <= prob[i])
  p3 <- p3data %>%
    ggplot(aes(x = p, y = Q)) +
    geom_line(col = "red") +
    geom_polygon(
      fill = "#ffbbbb",
      data = bind_rows(p3data, tibble(p = prob[i], Q = 0))
    ) +
    geom_line(data = tibble(p = c(prob[i], 1), Q = rep(qpactual, 2)), col = "red", linetype = 2) +
    # geom_point(data=tibble(p=prob[i],Q=qpactual), col='red') +
    labs(
      x = latex2exp::TeX("Probability p"),
      y = latex2exp::TeX("Quantile score $S_{p}$")
    ) +
    coord_cartesian(ylim = c(0, 11.6), xlim = c(0, 1), expand = FALSE, clip = "off")

  pdf(file = paste0("figs/qs", i, ".pdf"), width = 15 / 1.5, height = 8 / 1.5)
  print((plot_spacer() | p1) / (p3 | p2))
  crop::dev.off.crop()
}
```

\begin{textblock}{18}(0.3,1.4)
\animategraphics[width=14cm,height=20cm,keepaspectratio]{185}{figs/qs}{1}{185}
\end{textblock}

## Evaluating probabilistic forecasts

\begin{textblock}{18}(0.3,1.4)
\includegraphics[width=14cm,height=20cm,keepaspectratio]{figs/qs185}
\end{textblock}

\only<2>{\begin{textblock}{4}(2,7)\fontsize{11}{11}\sf
\textcolor{red}{Area = CRPS}
\end{textblock}}


## Evaluating probabilistic forecasts
\fontsize{9}{9}\sf

```{r fcaccuracya, dependson='holiday-fc'}
tourism_fc %>%
  accuracy(tourism, measures = list(MSE=MSE, CRPS=CRPS))
```

\vspace*{10cm}

## Evaluating probabilistic forecasts
\fontsize{9}{9}\sf

```{r fcaccuracy0, dependson='holiday-fc'}
tourism_fc %>%
  accuracy(tourism, measures = list(SS_MSE=skill_score(MSE), SS_CRPS=skill_score(CRPS)))
```

\vspace*{10cm}


## Evaluating probabilistic forecasts
\fontsize{9}{9}\sf

```{r fcaccuracy1, dependson='holiday-fc'}
tourism_fc %>%
  accuracy(tourism, measures = list(SS_MSE=skill_score(MSE), SS_CRPS=skill_score(CRPS))) %>%
  group_by(.model) %>%
  summarise(SS_MSE = mean(SS_MSE), SS_CRPS=mean(SS_CRPS)) %>%
  arrange(desc(SS_CRPS))
```

\vspace*{10cm}

## Tidyverts developers

\begin{textblock}{5.7}(0.4,1.4)
\begin{block}{Earo Wang}
\includegraphics[height=5.8cm]{earowang}
\end{block}
\end{textblock}

\begin{textblock}{5.7}(6.75,1.4)
\begin{block}{Mitchell O'Hara-Wild}
\includegraphics[height=5.8cm]{mitch2}
\end{block}
\end{textblock}

## More information
\fontsize{18}{20}\sf

 * Slides and papers: **robjhyndman.com**
 * Packages: **tidyverts.org**
 * Forecasting textbook using tidyverts package: **OTexts.com/fpp3**

\begin{textblock}{8}(7.6,4.8)
\begin{alertblock}{Find me at \dots}
\href{https://twitter.com/robjhyndman}{\faicon{twitter} @robjhyndman}

\href{https://github.com/robjhyndman}{\faicon{github} @robjhyndman}

\href{https://robjhyndman.com}{\faicon{home} robjhyndman.com}

\href{mailto:rob.hyndman@monash.edu}{\faicon{envelope} rob.hyndman@monash.edu}
\end{alertblock}
\end{textblock}
\vspace*{10cm}
